---
title: "RNA_setup"
author: "Lorenzo.MCM"
date: "2024-06-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Initial Setup

## Installing and loading packages.

```{r eval=FALSE}

install.packages(c('dplyr','biomaRt','readxl','DESeq2'), repos = "http://cran.us.r-project.org" )
```

```{r eval=FALSE}
BiocManager::install(c('DESeq2', 'AnnotationDbi', 'org.Mm.eg.db', 'apeglm', 'ReportingTools', 'clusterProfiler'))
```

```{r message=FALSE}
library(readxl)
library(DESeq2)
library(vsn)
library(Glimma)
library(regionReport)
library(clusterProfiler)
library(org.Mm.eg.db)
library(AnnotationDbi)
library(biomaRt)
library(stringr)
library(dplyr)
```

## Importing data and tidying up resulting data frame so rows are ID'd by Gene Symbol.NAs omitted for comparison with Gene list

```{r message=FALSE, warning=FALSE}

# Setting Working Directory to the folder which contains this .Rmd file

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# Importing main reads file

data = as.data.frame(read_excel("Reads_genes.xlsx"))

# Keeping only the columns with MGI, RefSeq ID and Gene symbols + raw counts, omitting NAs 

reads = data %>% select(MGI, Gene_ID, Gene_Symbol, contains("Count"))

reads = as.data.frame(reads)

reads = na.omit(reads)

```

## Importing panther-db list of RBPs and tidying it up (browsed for RNA Metabolism Proteins and filtered for Mus musculus, exported to and edited in excel so the only row was MGI IDs, removed genes that did not have MGI IDs)

```{r message=FALSE, warning=FALSE}

# Importing list of RBPs, omitting NAs

#mainPanther = as.data.frame(read_excel('rna_met.xlsx', col_names = FALSE))

mainPanther = as.data.frame(read_excel('Panther-RBPs.xlsx', col_names = FALSE))

mainPanther = mainPanther[!is.na(mainPanther)]

# Filtering the reads object so it only contains genes whose MGI is also present in the list of RBPs, resulting in a final "rbps" object 

rbps = reads[reads$MGI %in% mainPanther,]


```

## Naming conditions and final set up of rbps object

```{r message=FALSE, warning=FALSE}

# Naming rows in the rbps object after their MGI ID and removing all columns except those with the counts for each condition

rownames(rbps) = rbps$MGI

rbps = rbps[,-c(1,2,3)]

# Naming raw counts columns after their conditions, based on an external file

colnames(rbps) = (read_excel("conditions.xlsx", col_names = FALSE))[1,]
              
```

# Setting up for DESeq 
## Intermediate objects for selecting conditions for comparisons through DESeq

```{r message=FALSE, warning=FALSE}

#Splitting conditions for line and sex
rbpsF_WT = rbps %>% select(contains("FWT"))
rbpsM_WT = rbps %>% select(contains("MWT"))
rbpsF_TG = rbps %>% select(contains("F36"), contains("F41"))
rbpsM_TG = rbps %>% select(contains("M36"), contains("M41"))
rbpsF_36 = rbps %>% select(contains("F36"))
rbpsM_36 = rbps %>% select(contains("M36"))
rbpsF_41 = rbps %>% select(contains("F41"))
rbpsM_41 = rbps %>% select(contains("M41"))

#Final split into timepoints
T1F_41 = rbpsF_41[,grep("^1",names(rbpsF_41), value = TRUE)]
T1M_41 = rbpsM_41[,grep("^1",names(rbpsM_41), value = TRUE)]
T1F_WT = rbpsF_WT[,grep("^1",names(rbpsF_WT), value = TRUE)]
T1M_WT = rbpsM_WT[,grep("^1",names(rbpsM_WT), value = TRUE)]
T1F_36 = rbpsF_36[,grep("^1",names(rbpsF_36), value = TRUE)]
T1M_36 = rbpsM_36[,grep("^1",names(rbpsM_36), value = TRUE)]
T1F_TG = rbpsF_TG[,grep("^1",names(rbpsF_TG), value = TRUE)]
T1M_TG = rbpsM_TG[,grep("^1",names(rbpsM_TG), value = TRUE)]

T2F_41 = rbpsF_41[,grep("^2",names(rbpsF_41), value = TRUE)]
T2M_41 = rbpsM_41[,grep("^2",names(rbpsM_41), value = TRUE)]
T2F_WT = rbpsF_WT[,grep("^2",names(rbpsF_WT), value = TRUE)]
T2M_WT = rbpsM_WT[,grep("^2",names(rbpsM_WT), value = TRUE)]
T2F_36 = rbpsF_36[,grep("^2",names(rbpsF_36), value = TRUE)]
T2M_36 = rbpsM_36[,grep("^2",names(rbpsM_36), value = TRUE)]
T2F_TG = rbpsF_TG[,grep("^2",names(rbpsF_TG), value = TRUE)]
T2M_TG = rbpsM_TG[,grep("^2",names(rbpsM_TG), value = TRUE)]

T3F_41 = rbpsF_41[,grep("^3",names(rbpsF_41), value = TRUE)]
T3M_41 = rbpsM_41[,grep("^3",names(rbpsM_41), value = TRUE)]
T3F_WT = rbpsF_WT[,grep("^3",names(rbpsF_WT), value = TRUE)]
T3M_WT = rbpsM_WT[,grep("^3",names(rbpsM_WT), value = TRUE)]
T3F_36 = rbpsF_36[,grep("^3",names(rbpsF_36), value = TRUE)]
T3M_36 = rbpsM_36[,grep("^3",names(rbpsM_36), value = TRUE)]
T3F_TG = rbpsF_TG[,grep("^3",names(rbpsF_TG), value = TRUE)]
T3M_TG = rbpsM_TG[,grep("^3",names(rbpsM_TG), value = TRUE)]

#Binding final objects for comparison between different conditions (obj names = Timepoint-Sex-Line-conditions)
T1F_WT36 = bind_cols(T1F_WT, T1F_36)
T1M_WT36 = bind_cols(T1M_WT, T1M_36)
T1F_WT41 = bind_cols(T1F_WT, T1F_41)
T1M_WT41 = bind_cols(T1M_WT, T1M_41)
T1F_3641 = bind_cols(T1F_36, T1F_41)
T1M_3641 = bind_cols(T1M_36, T1M_41)
T1F_WTG = bind_cols(T1F_WT, T1F_TG)
T1M_WTG = bind_cols(T1M_WT, T1M_TG)
  
T2F_WT36 = bind_cols(T2F_WT, T2F_36)
T2M_WT36 = bind_cols(T2M_WT, T2M_36)
T2F_WT41 = bind_cols(T2F_WT, T2F_41)
T2M_WT41 = bind_cols(T2M_WT, T2M_41)
T2F_3641 = bind_cols(T2F_36, T2F_41)
T2M_3641 = bind_cols(T2M_36, T2M_41)
T2F_WTG = bind_cols(T2F_WT, T2F_TG)
T2M_WTG = bind_cols(T2M_WT, T2M_TG)

  
T3F_WT41 = bind_cols(T3F_WT, T3F_41)
T3M_WT41 = bind_cols(T3M_WT, T3M_41)
T3F_WT36 = bind_cols(T3F_WT, T3F_36)
T3M_WT36 = bind_cols(T3M_WT, T3M_36)
T3F_3641 = bind_cols(T3F_36, T3F_41)
T3M_3641 = bind_cols(T3M_36, T3M_41)
T3F_WTG = bind_cols(T3F_WT, T3F_TG)
T3M_WTG = bind_cols(T3M_WT, T3M_TG)

```

## setting up dfs for DESeq design 
```{r}
#timepoint lists
tpF_WT = c("1","1","1","2","2","2","3","3","3")
tpM_WT = c("1","1","1","2","2","2","3","3","3")
tpF_TG = c("1","1","1","1","2","2","2","3","1","1","1","2","2","2","3","3")
tpM_TG = c("1","1","2","2","2","3","3","3","3","3","1","1","1","2","2","2","3","3","3","3")

tpF_36 = c("1","1","1","1","2","2","2","3")
tpM_36 = c("1","1","2","2","2","3","3","3","3","3")

tpF_41 = c("1","1","1","2","2","2","3","3")
tpM_41 = c("1","1","1","2","2","2","3","3","3","3")

#line list

l1F_3641 = c("36","36","36","36","41","41","41")
l1M_3641 = c("36","36","41","41","41")

l1F_WTG = c("WT","WT","WT","TG","TG","TG","TG","TG","TG","TG")
l1M_WTG = c("WT","WT","WT","TG","TG","TG","TG","TG")

l1F_WT36 = c("WT","WT","WT","36","36","36","36")
l1M_WT36 = c("WT","WT","WT","36","36")

l1F_WT41 = c("WT","WT","WT","41","41","41")
l1M_WT41 = c("WT","WT","WT","41","41","41")

l2F_WTG = c("WT","WT","WT","TG","TG","TG","TG","TG","TG")
l2M_WTG = c("WT","WT","WT","TG","TG","TG","TG","TG","TG")

l2F_WT36 = c("WT","WT","WT","36","36","36")
l2M_WT36 = c("WT","WT","WT","36","36","36")

l2F_WT41 = c("WT","WT","WT","41","41","41")
l2M_WT41 = c("WT","WT","WT","41","41","41")

l2F_3641 = c("36","36","36","41","41","41")
l2M_3641 = c("36","36","36","41","41","41")

l3F_WTG = c("WT","WT","WT","TG","TG","TG")
l3M_WTG = c("WT","WT","WT","TG","TG","TG","TG","TG","TG","TG","TG","TG")

l3F_WT36 = c("WT","WT","WT","36")
l3M_WT36 = c("WT","WT","WT","36","36","36","36","36")

l3F_WT41 = c("WT","WT","WT","41","41")
l3M_WT41 = c("WT","WT","WT","41","41","41","41")

l3F_3641 = c("36","41","41")
l3M_3641 = c("36","36","36","36","36","41","41","41", "41")

# Binding timepoints and line lists for each condition

c1F_WTG = as.data.frame(cbind(colnames(T1F_WTG),l1F_WTG))
c1M_WTG = as.data.frame(cbind(colnames(T1M_WTG),l1M_WTG))

c1F_WT36 = as.data.frame(cbind(colnames(T1F_WT36),l1F_WT36))
c1M_WT36 = as.data.frame(cbind(colnames(T1M_WT36),l1M_WT36))

c1F_WT41 = as.data.frame(cbind(colnames(T1F_WT41),l1F_WT41))
c1M_WT41 = as.data.frame(cbind(colnames(T1M_WT41),l1M_WT41))

c1F_3641 = as.data.frame(cbind(colnames(T1F_3641),l1F_3641))
c1M_3641 = as.data.frame(cbind(colnames(T1M_3641),l1M_3641))

c2F_WTG = as.data.frame(cbind(colnames(T2F_WTG),l2F_WTG))
c2M_WTG = as.data.frame(cbind(colnames(T2M_WTG),l2M_WTG))

c2F_WT36 = as.data.frame(cbind(colnames(T2F_WT36),l2F_WT36))
c2M_WT36 = as.data.frame(cbind(colnames(T2M_WT36),l2M_WT36))

c2F_WT41 = as.data.frame(cbind(colnames(T2F_WT41),l2F_WT41))
c2M_WT41 = as.data.frame(cbind(colnames(T2M_WT41),l2M_WT41))

c2F_3641 = as.data.frame(cbind(colnames(T2F_3641),l2F_3641))
c2M_3641 = as.data.frame(cbind(colnames(T2M_3641),l2M_3641))

c3F_WTG = as.data.frame(cbind(colnames(T3F_WTG),l3F_WTG))
c3M_WTG = as.data.frame(cbind(colnames(T3M_WTG),l3M_WTG))

c3F_WT36 = as.data.frame(cbind(colnames(T3F_WT36),l3F_WT36))
c3M_WT36 = as.data.frame(cbind(colnames(T3M_WT36),l3M_WT36))

c3F_WT41 = as.data.frame(cbind(colnames(T3F_WT41),l3F_WT41))
c3M_WT41 = as.data.frame(cbind(colnames(T3M_WT41),l3M_WT41))

c3F_3641 = as.data.frame(cbind(colnames(T3F_3641),l3F_3641))
c3M_3641 = as.data.frame(cbind(colnames(T3M_3641),l3M_3641))

cF_WT = as.data.frame(cbind(colnames(rbpsF_WT),tpF_WT))
cM_WT = as.data.frame(cbind(colnames(rbpsM_WT),tpM_WT))

cF_TG = as.data.frame(cbind(colnames(rbpsF_TG),tpF_TG))
cM_TG = as.data.frame(cbind(colnames(rbpsM_TG),tpM_TG))

cF_36 = as.data.frame(cbind(colnames(rbpsF_36),tpF_36))
cM_36 = as.data.frame(cbind(colnames(rbpsM_36),tpM_36))

cF_41 = as.data.frame(cbind(colnames(rbpsF_41), tpF_41))
cM_41 = as.data.frame(cbind(colnames(rbpsM_41), tpM_41))


```

## Conversion into DESeqDataSets

```{r message=FALSE, warning=FALSE, include=FALSE}

dds1F_WTG = DESeqDataSetFromMatrix(countData = T1F_WTG,colData = c1F_WTG,
                                  design = ~l1F_WTG)
dds1M_WTG = DESeqDataSetFromMatrix(countData = T1M_WTG,colData = c1M_WTG,
                                  design = ~l1M_WTG)

dds1F_WT36 = DESeqDataSetFromMatrix(countData = T1F_WT36, colData = c1F_WT36,
                                 design = ~l1F_WT36)
dds1M_WT36 = DESeqDataSetFromMatrix(countData = T1M_WT36, colData = c1M_WT36,
                                 design = ~l1M_WT36)

dds1F_WT41 = DESeqDataSetFromMatrix(countData = T1F_WT41,colData = c1F_WT41,
                                 design = ~l1F_WT41)
dds1M_WT41 = DESeqDataSetFromMatrix(countData = T1M_WT41,colData = c1M_WT41,
                                 design = ~l1M_WT41)

dds1F_3641 = DESeqDataSetFromMatrix(countData = T1F_3641,colData = c1F_3641,
                                    design = ~l1F_3641)
dds1M_3641 = DESeqDataSetFromMatrix(countData = T1M_3641,colData = c1M_3641,
                                    design = ~l1M_3641)

dds2F_WTG = DESeqDataSetFromMatrix(countData = T2F_WTG,colData = c2F_WTG,
                                  design = ~l2F_WTG)
dds2M_WTG = DESeqDataSetFromMatrix(countData = T2M_WTG,colData = c2M_WTG,
                                  design = ~l2M_WTG)

dds2F_WT36 = DESeqDataSetFromMatrix(countData = T2F_WT36,colData = c2F_WT36,
                                 design = ~l2F_WT36)
dds2M_WT36 = DESeqDataSetFromMatrix(countData = T2M_WT36,colData = c2M_WT36,
                                 design = ~l2M_WT36)

dds2F_WT41 = DESeqDataSetFromMatrix(countData = T2F_WT41,colData = c2F_WT41,
                                 design = ~l2F_WT41)
dds2M_WT41 = DESeqDataSetFromMatrix(countData = T2M_WT41,colData = c2M_WT41,
                                 design = ~l2M_WT41)

dds2F_3641 = DESeqDataSetFromMatrix(countData = T2F_3641,colData = c2F_3641,
                                    design = ~l2F_3641)
dds2M_3641 = DESeqDataSetFromMatrix(countData = T2M_3641,colData = c2M_3641,
                                    design = ~l2M_3641)

dds3F_WTG = DESeqDataSetFromMatrix(countData = T3F_WTG,colData = c3F_WTG,
                                  design = ~l3F_WTG)
dds3M_WTG = DESeqDataSetFromMatrix(countData = T3M_WTG,colData = c3M_WTG,
                                  design = ~l3M_WTG)

dds3F_WT36 = DESeqDataSetFromMatrix(countData = T3F_WT36,colData = c3F_WT36,
                                 design = ~l3F_WT36)
dds3M_WT36 = DESeqDataSetFromMatrix(countData = T3M_WT36,colData = c3M_WT36,
                                 design = ~l3M_WT36)

dds3F_WT41 = DESeqDataSetFromMatrix(countData = T3F_WT41,colData = c3F_WT41,
                                 design = ~l3F_WT41)
dds3M_WT41 = DESeqDataSetFromMatrix(countData = T3M_WT41, colData = c3M_WT41,
                                 design = ~l3M_WT41)

dds3F_3641 = DESeqDataSetFromMatrix(countData = T3F_3641,colData = c3F_3641,
                                    design = ~l3F_3641)
dds3M_3641 = DESeqDataSetFromMatrix(countData = T3M_3641,colData = c3M_3641,
                                    design = ~l3M_3641)

ddsF_WT = DESeqDataSetFromMatrix(countData = rbpsF_WT,colData = cF_WT,
                                design = ~tpF_WT)
ddsM_WT = DESeqDataSetFromMatrix(countData = rbpsM_WT,colData = cM_WT,
                                design = ~tpM_WT)

ddsF_TG = DESeqDataSetFromMatrix(countData = rbpsF_TG,colData = cF_TG,
                                design = ~tpF_TG)
ddsM_TG = DESeqDataSetFromMatrix(countData = rbpsM_TG,colData = cM_TG,
                                design = ~tpM_TG)

ddsF_36 = DESeqDataSetFromMatrix(countData = rbpsF_36,colData = cF_36,
                                design = ~tpF_36)
ddsM_36 = DESeqDataSetFromMatrix(countData = rbpsM_36,colData = cM_36,
                                design = ~tpM_36)

ddsF_41 = DESeqDataSetFromMatrix(countData = rbpsF_41,colData = cF_41,
                                design = ~tpF_41)
ddsM_41 = DESeqDataSetFromMatrix(countData = rbpsM_41,colData = cM_41,
                                design = ~tpM_41)
```

## Running DESeq analysis

```{r message=FALSE}

dds1F_WTG = DESeq(dds1F_WTG)
dds1M_WTG = DESeq(dds1M_WTG)

dds1F_WT36 = DESeq(dds1F_WT36)
dds1M_WT36 = DESeq(dds1M_WT36)

dds1F_WT41 = DESeq(dds1F_WT41)
dds1M_WT41 = DESeq(dds1M_WT41)

dds1F_3641 = DESeq(dds1F_3641)
dds1M_3641 = DESeq(dds1M_3641)

dds2F_WTG = DESeq(dds2F_WTG)
dds2M_WTG = DESeq(dds2M_WTG)

dds2F_WT36 = DESeq(dds2F_WT36)
dds2M_WT36 = DESeq(dds2M_WT36)

dds2F_WT41 = DESeq(dds2F_WT41)
dds2M_WT41 = DESeq(dds2M_WT41)

dds2F_3641 = DESeq(dds2F_3641)
dds2M_3641 = DESeq(dds2M_3641)

dds3F_WTG = DESeq(dds3F_WTG)
dds3M_WTG = DESeq(dds3M_WTG)

dds3F_WT36 = DESeq(dds3F_WT36)
dds3M_WT36 = DESeq(dds3M_WT36)

dds3F_WT41 = DESeq(dds3F_WT41)
dds3M_WT41 = DESeq(dds3M_WT41)

dds3F_3641 = DESeq(dds3F_3641)
dds3M_3641 = DESeq(dds3M_3641)

ddsF_WT = DESeq(ddsF_WT)
ddsM_WT = DESeq(ddsM_WT)

ddsF_TG = DESeq(ddsF_TG)
ddsM_TG = DESeq(ddsM_TG)

ddsF_36 = DESeq(ddsF_36)
ddsM_36 = DESeq(ddsM_36)

ddsF_41 = DESeq(ddsF_41)
ddsM_41 = DESeq(ddsM_41)
```

# Preparing readable output

## Converting to result objects

```{r}

res1F_WTG = results(dds1F_WTG, contrast = c("l1F_WTG", "TG", "WT"))
res1M_WTG = results(dds1M_WTG, contrast = c("l1M_WTG", "TG", "WT"))

res1F_WT36 = results(dds1F_WT36, contrast = c("l1F_WT36", "36", "WT"))
res1M_WT36 = results(dds1M_WT36, contrast = c("l1M_WT36", "36", "WT"))

res1F_WT41 = results(dds1F_WT41, contrast = c("l1F_WT41", "41", "WT"))
res1M_WT41 = results(dds1M_WT41, contrast = c("l1M_WT41", "41", "WT"))

res1F_3641 = results(dds1F_3641, contrast = c("l1F_3641", "41", "36"))
res1M_3641 = results(dds1M_3641, contrast = c("l1M_3641", "41", "36"))

res2F_WTG = results(dds2F_WTG, contrast = c("l2F_WTG", "TG", "WT"))
res2M_WTG = results(dds2M_WTG, contrast = c("l2M_WTG", "TG", "WT"))

res2F_WT36 = results(dds2F_WT36, contrast = c("l2F_WT36", "36", "WT"))
res2M_WT36 = results(dds2M_WT36, contrast = c("l2M_WT36", "36", "WT"))

res2F_WT41 = results(dds2F_WT41, contrast = c("l2F_WT41", "41", "WT"))
res2M_WT41 = results(dds2M_WT41, contrast = c("l2M_WT41", "41", "WT"))

res2F_3641 = results(dds2F_3641, contrast = c("l2F_3641", "41", "36"))
res2M_3641 = results(dds2M_3641, contrast = c("l2M_3641", "41", "36"))

res3F_WTG = results(dds3F_WTG, contrast = c("l3F_WTG", "TG", "WT"))
res3M_WTG = results(dds3M_WTG, contrast = c("l3M_WTG", "TG", "WT"))

res3F_WT36 = results(dds3F_WT36, contrast = c("l3F_WT36", "36", "WT"))
res3M_WT36 = results(dds3M_WT36, contrast = c("l3M_WT36", "36", "WT"))

res3F_WT41 = results(dds3F_WT41, contrast = c("l3F_WT41", "41", "WT"))
res3M_WT41 = results(dds3M_WT41, contrast = c("l3M_WT41", "41", "WT"))

res3F_3641 = results(dds3F_3641, contrast = c("l3F_3641","41", "36"))
res3M_3641 = results(dds3M_3641, contrast = c("l3M_3641","41", "36"))
                    
resF_WT31 = results(ddsF_WT, contrast = c("tpF_WT","3", "1"))
resF_WT21 = results(ddsF_WT, contrast = c("tpF_WT","2", "1"))
resF_WT32 = results(ddsF_WT, contrast = c("tpF_WT","3", "2"))

resM_WT31 = results(ddsM_WT, contrast = c("tpM_WT","3","1"))
resM_WT21 = results(ddsM_WT, contrast = c("tpM_WT","2","1"))
resM_WT32 = results(ddsM_WT, contrast = c("tpM_WT","3","2"))

resF_TG31 = results(ddsF_TG, contrast = c("tpF_TG","3","1"))
resF_TG21 = results(ddsF_TG, contrast = c("tpF_TG","2","1"))
resF_TG32 = results(ddsF_TG, contrast = c("tpF_TG","3","2"))

resM_TG31 = results(ddsM_TG, contrast = c("tpM_TG","3","1"))
resM_TG21 = results(ddsM_TG, contrast = c("tpM_TG","2","1"))
resM_TG32 = results(ddsM_TG, contrast = c("tpM_TG","3","2"))

resF_36_31 = results(ddsF_36, contrast = c("tpF_36","3","1"))
resF_36_21 = results(ddsF_36, contrast = c("tpF_36","2","1"))
resF_36_32 = results(ddsF_36, contrast = c("tpF_36","3","2"))

resM_36_31 = results(ddsM_36, contrast = c("tpM_36","3","1"))
resM_36_21 = results(ddsM_36, contrast = c("tpM_36","2","1"))
resM_36_32 = results(ddsM_36, contrast = c("tpM_36","3","2"))

resF_41_31 = results(ddsF_41, contrast = c("tpF_41","3","1"))
resF_41_21 = results(ddsF_41, contrast = c("tpF_41","2","1"))
resF_41_32 = results(ddsF_41, contrast = c("tpF_41","3","2"))

resM_41_31 = results(ddsM_41, contrast = c("tpM_41","3","1"))
resM_41_21 = results(ddsM_41, contrast = c("tpM_41","2","1"))
resM_41_32 = results(ddsM_41, contrast = c("tpM_41","3","2"))

```

## Converting results to dfs, adding MGI,
entrez and Symbol as columns 

```{r message=FALSE}

# Making a list out of all results objects

allObj = ls()  

reObj = allObj[grep("^res", allObj)]

reList = mget(reObj)

# Adding back MGI, RefSeq ID and symbol columns to all results objects

reList = lapply(reList, function(res) {

  df = as.data.frame(res)

  df$MGI = rownames(df)
  
  df = merge(df, reads[, c("MGI", "Gene_ID", "Gene_Symbol")], by = "MGI", all.x = TRUE)
  
  rownames(df) <- NULL
  
  return(df)
})

```

## Function for selecting genes according to LFC, padj value and key

```{r}

pdegenes = function(x, lfcThreshold = 2, padjThreshold = 0.05, max = 20, key = "Gene_Symbol") {
  
  # Filter rows based on log2FoldChange and padj
  filtered_x <- x[(x$log2FoldChange > lfcThreshold) & (x$padj < padjThreshold) & !is.na(x$padj), ]
  
  # Sort the filtered rows by LFC value
  sorted_x <- filtered_x[order(filtered_x$log2FoldChange), ]
  
  # Handle max condition. If a number, returned DF has that many lines
  if (!is.null(max)) {
    sorted_x <- head(sorted_x, max)
  }
  
  return(sorted_x[[key]])
}

ndegenes = function(x, lfcThreshold = -2, padjThreshold = 0.05, max = 20, key = "Gene_Symbol") {

  filtered_x <- x[(x$log2FoldChange < lfcThreshold) & (x$padj < padjThreshold) & !is.na(x$padj), ]
  
  sorted_x <- filtered_x[order(filtered_x$log2FoldChange), ]
  
  if (!is.null(max)) {
    sorted_x <- head(sorted_x, max)
  }
  
  return(sorted_x[[key]])
}

degenes = function(x, lfcThreshold = 2, padjThreshold = 0.05, max = 20, key = "Gene_Symbol") {
  
  filtered_x <- x[(x$log2FoldChange > lfcThreshold | x$log2FoldChange < -lfcThreshold) & (x$padj < padjThreshold) & !is.na(x$padj),]
  
  sorted_x <- filtered_x[order(filtered_x$log2FoldChange), ]
  
  if (!is.null(max)) {
    sorted_x <- head(sorted_x, max)
  }
  
  return(sorted_x[[key]])
}

```

## Applying to all results

```{r}

genespAll = lapply(reList, pdegenes, lfcThreshold = 0.5, padjThreshold = 0.05, max = NULL)

genesnAll = lapply(reList, ndegenes, lfcThreshold = -0.5, padjThreshold = 0.05, max = NULL)

genesAll = lapply(reList, degenes, lfcThreshold = 0.5, padjThreshold = 0.05, max = NULL)

```

## Outputting as csv for easy consulting

```{r message=FALSE, warning=FALSE}

outputTidy = function(x, title) {

  max_length = max(sapply(x, length))
  
  padded_lists = lapply(x, function(y) {
    length(y) = max_length  
    return(y)
  })
  
  # Convert the list of lists into a data frame
  df = as.data.frame(padded_lists)
  
  # Step 3: Write the data frame to a CSV file
  write.csv(df, title, row.names = FALSE)
  
}

outputTidy(genesAll, "DEG_rbps.csv")
outputTidy(genespAll, "DEG_P_rbps.csv")
outputTidy(genesnAll, "DEG_N_rbps.csv")

#When opening the csv files in excel, navigate to data tab and click on "Text to columns" when selecting the entire first column, select "Delimited", add "comma" as a separator and format all columns as text. 


```

## Purkinje Cell Markers

```{r}
PC_markers = read_excel("pc_markers.xlsx",col_names = FALSE)

res1F_WT36[reList$res1F_WT36$MGI %in% PC_markers$...1,]

pc_mark = function(x){
  
  # Ensure 'MGI' column exists in the DESeq2 results object (x)
  if (!"MGI" %in% colnames(x)) {
    stop("Error: 'MGI' column not found in the DESeq2 results object")
  }
  
  # Subset 'x' where 'MGI' is present in the PC_markers column
  x_subset = x[x$MGI %in% PC_markers[[1]], ]
  
  x_subset = (x_subset[(x_subset$log2FoldChange > 0.5 | x_subset$log2FoldChange < -0.5) & x_subset$padj < 0.05,])
  
  x_subset = na.omit(x_subset)
  
  return(x_subset)
}
```

```{r}
pc_mark(reList$res1F_WT36)
pc_mark(reList$res1M_WT36)

pc_mark(reList$res2F_WT36)
pc_mark(reList$res2M_WT36)

pc_mark(reList$res3F_WT36)
pc_mark(reList$res3M_WT36)

pc_mark(reList$res1F_WT41)
pc_mark(reList$res1M_WT41)

pc_mark(reList$res2F_WT41)
pc_mark(reList$res2M_WT41)

pc_mark(reList$res3F_WT41)
pc_mark(reList$res3M_WT41)

```

