---
title: "GO_Analysis_RBPs"
author: "Lorenzo.MCM"
date: "2024-07-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Initial Setup

## Installing and loading packages.

```{r eval=FALSE}

install.packages(c('dplyr','readxl','DESeq2'), repos = "http://cran.us.r-project.org" )
```

```{r eval=FALSE}
BiocManager::install(c('DESeq2', 'AnnotationDbi', 'vsn', 'org.Mm.eg.db', 'apeglm', 'select'))
```

```{r message=FALSE}
library(readxl)
library(DESeq2)
library(ggplot2)
library(gplots)
library(pheatmap)
library(vsn)
library(Glimma)
library(regionReport)
library(clusterProfiler)
library(org.Mm.eg.db)
library(AnnotationDbi)
library(dplyr)
```

## Importing data and tidying up resulting data frame so rows are ID'd by Gene Symbol.NAs omitted for comparison with panther-db

```{r message=FALSE, warning=FALSE}

# Setting Working Directory (WD for easier output management)

#setwd("your/file/path")
setwd("C:/Users/Lorenzo/Documents/MESTRADO/R/RNAbps")

# Importing main reads file

data = as.data.frame(read_excel("Reads_genes.xlsx"))

# Keeping only the columns with MGI, RefSeq ID and Gene symbols + raw counts, omitting NAs 

reads = data %>% select(MGI, Gene_ID, Gene_Symbol, contains("Count"))

reads = as.data.frame(reads)

reads = na.omit(reads)

```

## Importing panther-db list of RBPs and tidying it up (browsed for RNA Metabolism Proteins and filtered for Mus musculus, exported to and edited in excel so the only row was MGI IDs, removed genes that did not have MGI IDs)

```{r message=FALSE, warning=FALSE}

# Importing list of RBPs, omitting NAs

mainPanther = as.data.frame(read_excel('Panther-RNA-ALL.xlsm', col_names = FALSE))

mainPanther = mainPanther[!is.na(mainPanther)]

# Filtering the reads object so it only contains genes whose MGI is also present in the list of RBPs, resulting in a final "rbps" object 
#(alternatively comment this first line and uncomment the last one to analyze all genes)

#comment this for analyzing all genes (uncomment last line)
rbps = reads[reads$MGI %in% mainPanther,]

#comment this for analyzing only rbps (uncomment previous line) 
#rbps = reads

```

## Naming conditions and final set up of rbps object

```{r message=FALSE, warning=FALSE}

# Naming rows in the rbps object after their MGI ID and removing all columns except those with the counts for each condition

rownames(rbps) = rbps$MGI

rbps = rbps[,-c(1,2,3)]

# Naming raw counts columns after their conditions, based on an external file

colnames(rbps) = (read_excel("conditions.xlsx", col_names = FALSE))[1,]
              
```

# Setting up for DESeq 
## Intermediate objects for selecting conditions for comparisons through DESeq

```{r message=FALSE, warning=FALSE}

#Splitting conditions for line and sex
rbpsF_WT = rbps %>% select(contains("FWT"))
rbpsM_WT = rbps %>% select(contains("MWT"))
rbpsF_TG = rbps %>% select(contains("F36"), contains("F41"))
rbpsM_TG = rbps %>% select(contains("M36"), contains("M41"))
rbpsF_36 = rbps %>% select(contains("F36"))
rbpsM_36 = rbps %>% select(contains("M36"))
rbpsF_41 = rbps %>% select(contains("F41"))
rbpsM_41 = rbps %>% select(contains("M41"))

#Final split into timepoints
T1F_41 = rbpsF_41[,grep("^1",names(rbpsF_41), value = TRUE)]
T1M_41 = rbpsM_41[,grep("^1",names(rbpsM_41), value = TRUE)]
T1F_WT = rbpsF_WT[,grep("^1",names(rbpsF_WT), value = TRUE)]
T1M_WT = rbpsM_WT[,grep("^1",names(rbpsM_WT), value = TRUE)]
T1F_36 = rbpsF_36[,grep("^1",names(rbpsF_36), value = TRUE)]
T1M_36 = rbpsM_36[,grep("^1",names(rbpsM_36), value = TRUE)]
T1F_TG = rbpsF_TG[,grep("^1",names(rbpsF_TG), value = TRUE)]
T1M_TG = rbpsM_TG[,grep("^1",names(rbpsM_TG), value = TRUE)]

T2F_41 = rbpsF_41[,grep("^2",names(rbpsF_41), value = TRUE)]
T2M_41 = rbpsM_41[,grep("^2",names(rbpsM_41), value = TRUE)]
T2F_WT = rbpsF_WT[,grep("^2",names(rbpsF_WT), value = TRUE)]
T2M_WT = rbpsM_WT[,grep("^2",names(rbpsM_WT), value = TRUE)]
T2F_36 = rbpsF_36[,grep("^2",names(rbpsF_36), value = TRUE)]
T2M_36 = rbpsM_36[,grep("^2",names(rbpsM_36), value = TRUE)]
T2F_TG = rbpsF_TG[,grep("^2",names(rbpsF_TG), value = TRUE)]
T2M_TG = rbpsM_TG[,grep("^2",names(rbpsM_TG), value = TRUE)]

T3F_41 = rbpsF_41[,grep("^3",names(rbpsF_41), value = TRUE)]
T3M_41 = rbpsM_41[,grep("^3",names(rbpsM_41), value = TRUE)]
T3F_WT = rbpsF_WT[,grep("^3",names(rbpsF_WT), value = TRUE)]
T3M_WT = rbpsM_WT[,grep("^3",names(rbpsM_WT), value = TRUE)]
T3F_36 = rbpsF_36[,grep("^3",names(rbpsF_36), value = TRUE)]
T3M_36 = rbpsM_36[,grep("^3",names(rbpsM_36), value = TRUE)]
T3F_TG = rbpsF_TG[,grep("^3",names(rbpsF_TG), value = TRUE)]
T3M_TG = rbpsM_TG[,grep("^3",names(rbpsM_TG), value = TRUE)]

#Binding final objects for comparison between different conditions (obj names = Timepoint-Sex-Line-conditions)
T1F_WT36 = bind_cols(T1F_WT, T1F_36)
T1M_WT36 = bind_cols(T1M_WT, T1M_36)
T1F_WT41 = bind_cols(T1F_WT, T1F_41)
T1M_WT41 = bind_cols(T1M_WT, T1M_41)
T1F_3641 = bind_cols(T1F_36, T1F_41)
T1M_3641 = bind_cols(T1M_36, T1M_41)
T1F_WTG = bind_cols(T1F_WT, T1F_TG)
T1M_WTG = bind_cols(T1M_WT, T1M_TG)
  
T2F_WT36 = bind_cols(T2F_WT, T2F_36)
T2M_WT36 = bind_cols(T2M_WT, T2M_36)
T2F_WT41 = bind_cols(T2F_WT, T2F_41)
T2M_WT41 = bind_cols(T2M_WT, T2M_41)
T2F_3641 = bind_cols(T2F_36, T2F_41)
T2M_3641 = bind_cols(T2M_36, T2M_41)
T2F_WTG = bind_cols(T2F_WT, T2F_TG)
T2M_WTG = bind_cols(T2M_WT, T2M_TG)

  
T3F_WT41 = bind_cols(T3F_WT, T3F_41)
T3M_WT41 = bind_cols(T3M_WT, T3M_41)
T3F_WT36 = bind_cols(T3F_WT, T3F_36)
T3M_WT36 = bind_cols(T3M_WT, T3M_36)
T3F_3641 = bind_cols(T3F_36, T3F_41)
T3M_3641 = bind_cols(T3M_36, T3M_41)
T3F_WTG = bind_cols(T3F_WT, T3F_TG)
T3M_WTG = bind_cols(T3M_WT, T3M_TG)

```

## setting up dfs for DESeq design 
```{r}
#timepoint lists
tpF_WT = c("1","1","1","2","2","2","3","3","3")
tpM_WT = c("1","1","1","2","2","2","3","3","3")
tpF_TG = c("1","1","1","1","2","2","2","3","1","1","1","2","2","2","3","3")
tpM_TG = c("1","1","2","2","2","3","3","3","3","3","1","1","1","2","2","2","3","3","3","3")

tpF_36 = c("1","1","1","1","2","2","2","3")
tpM_36 = c("1","1","2","2","2","3","3","3","3","3")

tpF_41 = c("1","1","1","2","2","2","3","3")
tpM_41 = c("1","1","1","2","2","2","3","3","3","3")

#line list

l1F_3641 = c("36","36","36","36","41","41","41")
l1M_3641 = c("36","36","41","41","41")

l1F_WTG = c("WT","WT","WT","TG","TG","TG","TG","TG","TG","TG")
l1M_WTG = c("WT","WT","WT","TG","TG","TG","TG","TG")

l1F_WT36 = c("WT","WT","WT","36","36","36","36")
l1M_WT36 = c("WT","WT","WT","36","36")

l1F_WT41 = c("WT","WT","WT","41","41","41")
l1M_WT41 = c("WT","WT","WT","41","41","41")

l2F_WTG = c("WT","WT","WT","TG","TG","TG","TG","TG","TG")
l2M_WTG = c("WT","WT","WT","TG","TG","TG","TG","TG","TG")

l2F_WT36 = c("WT","WT","WT","36","36","36")
l2M_WT36 = c("WT","WT","WT","36","36","36")

l2F_WT41 = c("WT","WT","WT","41","41","41")
l2M_WT41 = c("WT","WT","WT","41","41","41")

l2F_3641 = c("36","36","36","41","41","41")
l2M_3641 = c("36","36","36","41","41","41")

l3F_WTG = c("WT","WT","WT","TG","TG","TG")
l3M_WTG = c("WT","WT","WT","TG","TG","TG","TG","TG","TG","TG","TG","TG")

l3F_WT36 = c("WT","WT","WT","36")
l3M_WT36 = c("WT","WT","WT","36","36","36","36","36")

l3F_WT41 = c("WT","WT","WT","41","41")
l3M_WT41 = c("WT","WT","WT","41","41","41","41")

l3F_3641 = c("36","41","41")
l3M_3641 = c("36","36","36","36","36","41","41","41", "41")

# Binding timepoints and line lists for each condition

c1F_WTG = as.data.frame(cbind(colnames(T1F_WTG),l1F_WTG))
c1M_WTG = as.data.frame(cbind(colnames(T1M_WTG),l1M_WTG))

c1F_WT36 = as.data.frame(cbind(colnames(T1F_WT36),l1F_WT36))
c1M_WT36 = as.data.frame(cbind(colnames(T1M_WT36),l1M_WT36))

c1F_WT41 = as.data.frame(cbind(colnames(T1F_WT41),l1F_WT41))
c1M_WT41 = as.data.frame(cbind(colnames(T1M_WT41),l1M_WT41))

c1F_3641 = as.data.frame(cbind(colnames(T1F_3641),l1F_3641))
c1M_3641 = as.data.frame(cbind(colnames(T1M_3641),l1M_3641))

c2F_WTG = as.data.frame(cbind(colnames(T2F_WTG),l2F_WTG))
c2M_WTG = as.data.frame(cbind(colnames(T2M_WTG),l2M_WTG))

c2F_WT36 = as.data.frame(cbind(colnames(T2F_WT36),l2F_WT36))
c2M_WT36 = as.data.frame(cbind(colnames(T2M_WT36),l2M_WT36))

c2F_WT41 = as.data.frame(cbind(colnames(T2F_WT41),l2F_WT41))
c2M_WT41 = as.data.frame(cbind(colnames(T2M_WT41),l2M_WT41))

c2F_3641 = as.data.frame(cbind(colnames(T2F_3641),l2F_3641))
c2M_3641 = as.data.frame(cbind(colnames(T2M_3641),l2M_3641))

c3F_WTG = as.data.frame(cbind(colnames(T3F_WTG),l3F_WTG))
c3M_WTG = as.data.frame(cbind(colnames(T3M_WTG),l3M_WTG))

c3F_WT36 = as.data.frame(cbind(colnames(T3F_WT36),l3F_WT36))
c3M_WT36 = as.data.frame(cbind(colnames(T3M_WT36),l3M_WT36))

c3F_WT41 = as.data.frame(cbind(colnames(T3F_WT41),l3F_WT41))
c3M_WT41 = as.data.frame(cbind(colnames(T3M_WT41),l3M_WT41))

c3F_3641 = as.data.frame(cbind(colnames(T3F_3641),l3F_3641))
c3M_3641 = as.data.frame(cbind(colnames(T3M_3641),l3M_3641))

cF_WT = as.data.frame(cbind(colnames(rbpsF_WT),tpF_WT))
cM_WT = as.data.frame(cbind(colnames(rbpsM_WT),tpM_WT))

cF_TG = as.data.frame(cbind(colnames(rbpsF_TG),tpF_TG))
cM_TG = as.data.frame(cbind(colnames(rbpsM_TG),tpM_TG))

cF_36 = as.data.frame(cbind(colnames(rbpsF_36),tpF_36))
cM_36 = as.data.frame(cbind(colnames(rbpsM_36),tpM_36))

cF_41 = as.data.frame(cbind(colnames(rbpsF_41), tpF_41))
cM_41 = as.data.frame(cbind(colnames(rbpsM_41), tpM_41))


```

## Conversion into DESeqDataSets

```{r message=FALSE, warning=FALSE, include=FALSE}

dds1F_WTG = DESeqDataSetFromMatrix(countData = T1F_WTG,colData = c1F_WTG,
                                  design = ~l1F_WTG)
dds1M_WTG = DESeqDataSetFromMatrix(countData = T1M_WTG,colData = c1M_WTG,
                                  design = ~l1M_WTG)

dds1F_WT36 = DESeqDataSetFromMatrix(countData = T1F_WT36, colData = c1F_WT36,
                                 design = ~l1F_WT36)
dds1M_WT36 = DESeqDataSetFromMatrix(countData = T1M_WT36, colData = c1M_WT36,
                                 design = ~l1M_WT36)

dds1F_WT41 = DESeqDataSetFromMatrix(countData = T1F_WT41,colData = c1F_WT41,
                                 design = ~l1F_WT41)
dds1M_WT41 = DESeqDataSetFromMatrix(countData = T1M_WT41,colData = c1M_WT41,
                                 design = ~l1M_WT41)

dds1F_3641 = DESeqDataSetFromMatrix(countData = T1F_3641,colData = c1F_3641,
                                    design = ~l1F_3641)
dds1M_3641 = DESeqDataSetFromMatrix(countData = T1M_3641,colData = c1M_3641,
                                    design = ~l1M_3641)

dds2F_WTG = DESeqDataSetFromMatrix(countData = T2F_WTG,colData = c2F_WTG,
                                  design = ~l2F_WTG)
dds2M_WTG = DESeqDataSetFromMatrix(countData = T2M_WTG,colData = c2M_WTG,
                                  design = ~l2M_WTG)

dds2F_WT36 = DESeqDataSetFromMatrix(countData = T2F_WT36,colData = c2F_WT36,
                                 design = ~l2F_WT36)
dds2M_WT36 = DESeqDataSetFromMatrix(countData = T2M_WT36,colData = c2M_WT36,
                                 design = ~l2M_WT36)

dds2F_WT41 = DESeqDataSetFromMatrix(countData = T2F_WT41,colData = c2F_WT41,
                                 design = ~l2F_WT41)
dds2M_WT41 = DESeqDataSetFromMatrix(countData = T2M_WT41,colData = c2M_WT41,
                                 design = ~l2M_WT41)

dds2F_3641 = DESeqDataSetFromMatrix(countData = T2F_3641,colData = c2F_3641,
                                    design = ~l2F_3641)
dds2M_3641 = DESeqDataSetFromMatrix(countData = T2M_3641,colData = c2M_3641,
                                    design = ~l2M_3641)

dds3F_WTG = DESeqDataSetFromMatrix(countData = T3F_WTG,colData = c3F_WTG,
                                  design = ~l3F_WTG)
dds3M_WTG = DESeqDataSetFromMatrix(countData = T3M_WTG,colData = c3M_WTG,
                                  design = ~l3M_WTG)

dds3F_WT36 = DESeqDataSetFromMatrix(countData = T3F_WT36,colData = c3F_WT36,
                                 design = ~l3F_WT36)
dds3M_WT36 = DESeqDataSetFromMatrix(countData = T3M_WT36,colData = c3M_WT36,
                                 design = ~l3M_WT36)

dds3F_WT41 = DESeqDataSetFromMatrix(countData = T3F_WT41,colData = c3F_WT41,
                                 design = ~l3F_WT41)
dds3M_WT41 = DESeqDataSetFromMatrix(countData = T3M_WT41, colData = c3M_WT41,
                                 design = ~l3M_WT41)

dds3F_3641 = DESeqDataSetFromMatrix(countData = T3F_3641,colData = c3F_3641,
                                    design = ~l3F_3641)
dds3M_3641 = DESeqDataSetFromMatrix(countData = T3M_3641,colData = c3M_3641,
                                    design = ~l3M_3641)

ddsF_WT = DESeqDataSetFromMatrix(countData = rbpsF_WT,colData = cF_WT,
                                design = ~tpF_WT)
ddsM_WT = DESeqDataSetFromMatrix(countData = rbpsM_WT,colData = cM_WT,
                                design = ~tpM_WT)

ddsF_TG = DESeqDataSetFromMatrix(countData = rbpsF_TG,colData = cF_TG,
                                design = ~tpF_TG)
ddsM_TG = DESeqDataSetFromMatrix(countData = rbpsM_TG,colData = cM_TG,
                                design = ~tpM_TG)

ddsF_36 = DESeqDataSetFromMatrix(countData = rbpsF_36,colData = cF_36,
                                design = ~tpF_36)
ddsM_36 = DESeqDataSetFromMatrix(countData = rbpsM_36,colData = cM_36,
                                design = ~tpM_36)

ddsF_41 = DESeqDataSetFromMatrix(countData = rbpsF_41,colData = cF_41,
                                design = ~tpF_41)
ddsM_41 = DESeqDataSetFromMatrix(countData = rbpsM_41,colData = cM_41,
                                design = ~tpM_41)
```

## Running DESeq

```{r message=FALSE, warning=FALSE}

dds1F_WTG = DESeq(dds1F_WTG)
dds1M_WTG = DESeq(dds1M_WTG)

dds1F_WT36 = DESeq(dds1F_WT36)
dds1M_WT36 = DESeq(dds1M_WT36)

dds1F_WT41 = DESeq(dds1F_WT41)
dds1M_WT41 = DESeq(dds1M_WT41)

dds1F_3641 = DESeq(dds1F_3641)
dds1M_3641 = DESeq(dds1M_3641)

dds2F_WTG = DESeq(dds2F_WTG)
dds2M_WTG = DESeq(dds2M_WTG)

dds2F_WT36 = DESeq(dds2F_WT36)
dds2M_WT36 = DESeq(dds2M_WT36)

dds2F_WT41 = DESeq(dds2F_WT41)
dds2M_WT41 = DESeq(dds2M_WT41)

dds2F_3641 = DESeq(dds2F_3641)
dds2M_3641 = DESeq(dds2M_3641)

dds3F_WTG = DESeq(dds3F_WTG)
dds3M_WTG = DESeq(dds3M_WTG)

dds3F_WT36 = DESeq(dds3F_WT36)
dds3M_WT36 = DESeq(dds3M_WT36)

dds3F_WT41 = DESeq(dds3F_WT41)
dds3M_WT41 = DESeq(dds3M_WT41)

dds3F_3641 = DESeq(dds3F_3641)
dds3M_3641 = DESeq(dds3M_3641)

ddsF_WT = DESeq(ddsF_WT)
ddsM_WT = DESeq(ddsM_WT)

ddsF_TG = DESeq(ddsF_TG)
ddsM_TG = DESeq(ddsM_TG)

ddsF_36 = DESeq(ddsF_36)
ddsM_36 = DESeq(ddsM_36)

ddsF_41 = DESeq(ddsF_41)
ddsM_41 = DESeq(ddsM_41)
```

# Reports on differentially expressed (Padj < 0.5) genes for easy visualization, commented to allow Knitting

```{r eval=FALSE, include=FALSE}
# Commented to allow Markdown knitting, must uncomment and open output in browser for visualization

#limmaVolcano(x = dds3F_WTG, counts = T3F_WTG, groups = c3F_WTG$l3F_WTG)

#DESeq2Report(dds = dds3F_WTG, project = "T3 Female TG v WT", intgroup = c("l3F_WTG"), output = "T3_F_WTG")
```

# Preparing readable output

## Converting to result objects

```{r}

res1F_WTG = results(dds1F_WTG, contrast = c("l1F_WTG", "TG", "WT"))
res1M_WTG = results(dds1M_WTG, contrast = c("l1M_WTG", "TG", "WT"))

res1F_WT36 = results(dds1F_WT36, contrast = c("l1F_WT36", "36", "WT"))
res1M_WT36 = results(dds1M_WT36, contrast = c("l1M_WT36", "36", "WT"))

res1F_WT41 = results(dds1F_WT41, contrast = c("l1F_WT41", "41", "WT"))
res1M_WT41 = results(dds1M_WT41, contrast = c("l1M_WT41", "41", "WT"))

res1F_3641 = results(dds1F_3641, contrast = c("l1F_3641", "41", "36"))
res1M_3641 = results(dds1M_3641, contrast = c("l1M_3641", "41", "36"))

res2F_WTG = results(dds2F_WTG, contrast = c("l2F_WTG", "TG", "WT"))
res2M_WTG = results(dds2M_WTG, contrast = c("l2M_WTG", "TG", "WT"))

res2F_WT36 = results(dds2F_WT36, contrast = c("l2F_WT36", "36", "WT"))
res2M_WT36 = results(dds2M_WT36, contrast = c("l2M_WT36", "36", "WT"))

res2F_WT41 = results(dds2F_WT41, contrast = c("l2F_WT41", "41", "WT"))
res2M_WT41 = results(dds2M_WT41, contrast = c("l2M_WT41", "41", "WT"))

res2F_3641 = results(dds2F_3641, contrast = c("l2F_3641", "41", "36"))
res2M_3641 = results(dds2M_3641, contrast = c("l2M_3641", "41", "36"))

res3F_WTG = results(dds3F_WTG, contrast = c("l3F_WTG", "TG", "WT"))
res3M_WTG = results(dds3M_WTG, contrast = c("l3M_WTG", "TG", "WT"))

res3F_WT36 = results(dds3F_WT36, contrast = c("l3F_WT36", "36", "WT"))
res3M_WT36 = results(dds3M_WT36, contrast = c("l3M_WT36", "36", "WT"))

res3F_WT41 = results(dds3F_WT41, contrast = c("l3F_WT41", "41", "WT"))
res3M_WT41 = results(dds3M_WT41, contrast = c("l3M_WT41", "41", "WT"))

res3F_3641 = results(dds3F_3641, contrast = c("l3F_3641","41", "36"))
res3M_3641 = results(dds3M_3641, contrast = c("l3M_3641","41", "36"))
                    
resF_WT31 = results(ddsF_WT, contrast = c("tpF_WT","3", "1"))
resF_WT21 = results(ddsF_WT, contrast = c("tpF_WT","2", "1"))
resF_WT32 = results(ddsF_WT, contrast = c("tpF_WT","3", "2"))

resM_WT31 = results(ddsM_WT, contrast = c("tpM_WT","3","1"))
resM_WT21 = results(ddsM_WT, contrast = c("tpM_WT","2","1"))
resM_WT32 = results(ddsM_WT, contrast = c("tpM_WT","3","2"))

resF_TG31 = results(ddsF_TG, contrast = c("tpF_TG","3","1"))
resF_TG21 = results(ddsF_TG, contrast = c("tpF_TG","2","1"))
resF_TG32 = results(ddsF_TG, contrast = c("tpF_TG","3","2"))

resM_TG31 = results(ddsM_TG, contrast = c("tpM_TG","3","1"))
resM_TG21 = results(ddsM_TG, contrast = c("tpM_TG","2","1"))
resM_TG32 = results(ddsM_TG, contrast = c("tpM_TG","3","2"))

resF_36_31 = results(ddsF_36, contrast = c("tpF_36","3","1"))
resF_36_21 = results(ddsF_36, contrast = c("tpF_36","2","1"))
resF_36_32 = results(ddsF_36, contrast = c("tpF_36","3","2"))

resM_36_31 = results(ddsM_36, contrast = c("tpM_36","3","1"))
resM_36_21 = results(ddsM_36, contrast = c("tpM_36","2","1"))
resM_36_32 = results(ddsM_36, contrast = c("tpM_36","3","2"))

resF_41_31 = results(ddsF_41, contrast = c("tpF_41","3","1"))
resF_41_21 = results(ddsF_41, contrast = c("tpF_41","2","1"))
resF_41_32 = results(ddsF_41, contrast = c("tpF_41","3","2"))

resM_41_31 = results(ddsM_41, contrast = c("tpM_41","3","1"))
resM_41_21 = results(ddsM_41, contrast = c("tpM_41","2","1"))
resM_41_32 = results(ddsM_41, contrast = c("tpM_41","3","2"))
```

# GO Analysis

```{r message=FALSE}
library(GO.db)
library(GOstats)
```
## Converting results to dfs, adding MGI,
entrez and Symbol as columns 

```{r message=FALSE}

# Making a list out of all results objects

allObj = ls()  

reObj = allObj[grep("^res", allObj)]

reList = mget(reObj)

# Adding back MGI, RefSeq ID and symbol columns to all results objects

reList = lapply(reList, function(res) {

  df = as.data.frame(res)

  df$MGI = rownames(df)
  
  df = merge(df, reads[, c("MGI", "Gene_ID", "Gene_Symbol")], by = "MGI", all.x = TRUE)
  
  rownames(df) <- NULL
  
  return(df)
})
```

```{r}
#Subsetting only genes with a padj below 0.05 from results
sigRes = lapply(reList, function(x) as.data.frame(subset(x, padj < 0.05)))

#Functions for GO analysis
go_PDE = function(x, lfcThreshold = 2, cutOff = 0.05, ont = "BP"){
  
  selectGenes = unique(x$Gene_ID[x$log2FoldChange > lfcThreshold])
  
  if(length(selectGenes) == 0){
    warning("No genes meet the log2 fold change threshold.")
    return(NULL)
  }  
  
  universe = unique(x$'Gene_ID')

  Params = new("GOHyperGParams",
                 geneIds = selectGenes,
                 universeGeneIds = universe,
                 annotation = "org.Mm.eg.db",
                 ontology = ont,
                 pvalueCutoff = cutOff,
                 conditional = FALSE,
                 testDirection = "over")
  
  return(hyperGTest(Params))
  
}
go_NDE = function(x, lfcThreshold = -2, cutOff = 0.05, ont = "BP"){
  
  selectGenes = unique(x$Gene_ID[x$log2FoldChange < lfcThreshold])
  
  if(length(selectGenes) == 0){
    warning("No genes meet the log2 fold change threshold.")
    return(NULL)
  }  
  
  universe = unique(x$'Gene_ID')

  Params = new("GOHyperGParams",
                 geneIds = selectGenes,
                 universeGeneIds = universe,
                 annotation = "org.Mm.eg.db",
                 ontology = ont,
                 pvalueCutoff = cutOff,
                 conditional = FALSE,
                 testDirection = "over")
  
  return(hyperGTest(Params))
  
}

allgo = function(x, lfcThreshold = 2, cutOff = 0.05, ont = "BP"){
  
  selectGenes = unique(x$Gene_ID[x$log2FoldChange > lfcThreshold | x$log2FoldChange < -lfcThreshold])
  
  if(length(selectGenes) == 0){
    warning("No genes meet the log2 fold change threshold.")
    return(NULL)
  }  
  
  universe <- unique(x$Gene_ID)
  
  Params = new("GOHyperGParams",
                 geneIds = selectGenes,
                 universeGeneIds = universe,
                 annotation = "org.Mm.eg.db",
                 ontology = ont,
                 pvalueCutoff = cutOff,
                 conditional = FALSE,
                 testDirection = "over")
    
  return(hyperGTest(Params))
}


```

## Comparing WT and TG mice (positively and negatively DE genes and then all) (these take ~5 minutes to complete per chunk, depending on your CPU)

### BP Analysis
```{r}
BPGOP1F_WTG = (summary(go_PDE(sigRes$res1F_WTG)))
BPGOP1M_WTG = (summary(go_PDE(sigRes$res1M_WTG)))
BPGON1F_WTG = (summary(go_NDE(sigRes$res1F_WTG)))
BPGON1M_WTG = (summary(go_NDE(sigRes$res1M_WTG)))

BPGOP2F_WTG = (summary(go_PDE(sigRes$res2F_WTG)))
BPGOP2M_WTG = (summary(go_PDE(sigRes$res2M_WTG)))
BPGON2F_WTG = (summary(go_NDE(sigRes$res2F_WTG)))
BPGON2M_WTG = (summary(go_NDE(sigRes$res2M_WTG)))

BPGOP3F_WTG = (summary(go_PDE(sigRes$res3F_WTG)))
BPGOP3M_WTG = (summary(go_PDE(sigRes$res3M_WTG)))
BPGON3F_WTG = (summary(go_NDE(sigRes$res3F_WTG)))
BPGON3M_WTG = (summary(go_NDE(sigRes$res3M_WTG)))
# P for POSITIVELY DGE
# N for NEGATIVELY DGE

BPGOP1F_WTG
BPGOP1M_WTG
```

```{r}
BPGON1F_WTG
BPGON1M_WTG
```

```{r}
BPGOP2F_WTG
BPGOP2M_WTG
```

```{r}
BPGON2F_WTG
BPGON2M_WTG
```

```{r}
BPGOP3F_WTG
BPGOP3M_WTG
```

```{r}
BPGON3F_WTG
BPGON3M_WTG
```

### CC Analysis

```{r}
CCGOP1F_WTG = summary(go_PDE(sigRes$res1F_WTG, ont = "CC"))
CCGOP1M_WTG = summary(go_PDE(sigRes$res1M_WTG, ont = "CC"))
CCGON1F_WTG = summary(go_NDE(sigRes$res1F_WTG, ont = "CC"))
CCGON1M_WTG = summary(go_NDE(sigRes$res1M_WTG, ont = "CC"))

CCGOP2F_WTG = summary(go_PDE(sigRes$res2F_WTG, ont = "CC"))
CCGOP2M_WTG = summary(go_PDE(sigRes$res2M_WTG, ont = "CC"))
CCGON2F_WTG = summary(go_NDE(sigRes$res2F_WTG, ont = "CC"))
CCGON2M_WTG = summary(go_NDE(sigRes$res2M_WTG, ont = "CC"))

CCGOP3F_WTG = summary(go_PDE(sigRes$res3F_WTG, ont = "CC"))
CCGOP3M_WTG = summary(go_PDE(sigRes$res3M_WTG, ont = "CC"))
CCGON3F_WTG = summary(go_NDE(sigRes$res3F_WTG, ont = "CC"))
CCGON3M_WTG = summary(go_NDE(sigRes$res3M_WTG, ont = "CC"))

CCGOP1F_WTG
CCGOP1M_WTG
```

```{r}
CCGON1F_WTG
CCGON1M_WTG
```


```{r}
CCGOP2F_WTG
CCGOP2M_WTG
```

```{r}
CCGON2F_WTG
CCGON2M_WTG
```

```{r}
CCGOP3F_WTG
CCGOP3M_WTG
```

```{r}
CCGON3F_WTG
CCGON3M_WTG
```

### MF Analysis

```{r}
MFGOP1F_WTG = summary(go_PDE(sigRes$res1F_WTG, ont = "MF"))
MFGOP1M_WTG = summary(go_PDE(sigRes$res1M_WTG, ont = "MF"))
MFGON1F_WTG = summary(go_NDE(sigRes$res1F_WTG, ont = "MF"))
MFGON1M_WTG = summary(go_NDE(sigRes$res1M_WTG, ont = "MF"))

MFGOP2F_WTG = summary(go_PDE(sigRes$res2F_WTG, ont = "MF"))
MFGOP2M_WTG = summary(go_PDE(sigRes$res2M_WTG, ont = "MF"))
MFGON2F_WTG = summary(go_NDE(sigRes$res2F_WTG, ont = "MF"))
MFGON2M_WTG = summary(go_NDE(sigRes$res2M_WTG, ont = "MF"))

MFGOP3F_WTG = summary(go_PDE(sigRes$res3F_WTG, ont = "MF"))
MFGOP3M_WTG = summary(go_PDE(sigRes$res3M_WTG, ont = "MF"))
MFGON3F_WTG = summary(go_NDE(sigRes$res3F_WTG, ont = "MF"))
MFGON3M_WTG = summary(go_NDE(sigRes$res3M_WTG, ont = "MF"))

MFGOP1F_WTG
MFGOP1M_WTG
```

```{r}
MFGON1F_WTG
MFGON1M_WTG
```


```{r}
MFGOP2F_WTG
MFGOP2M_WTG
```

```{r}
MFGON2F_WTG
MFGON2M_WTG
```

```{r}
MFGOP3F_WTG
MFGOP3M_WTG
```

```{r}
MFGON3F_WTG
MFGON3M_WTG
```

### BP PDE & NDE Analysis

```{r}
BPGO1F_WTG = (summary(allgo(sigRes$res1F_WTG)))
BPGO1M_WTG = (summary(allgo(sigRes$res1M_WTG)))
BPGO2F_WTG = (summary(allgo(sigRes$res2F_WTG)))
BPGO2M_WTG = (summary(allgo(sigRes$res2M_WTG)))
BPGO3F_WTG = (summary(allgo(sigRes$res3F_WTG)))
BPGO3M_WTG = (summary(allgo(sigRes$res3M_WTG)))


BPGO1F_WTG
BPGO1M_WTG
```

```{r}
BPGO2F_WTG
BPGO2M_WTG
```

```{r}
BPGO3F_WTG
BPGO3M_WTG
```

### CC PDE & NDE Analysis

```{r}
CCGO1F_WTG = summary(allgo(sigRes$res1F_WTG, ont = "CC"))
CCGO1M_WTG = summary(allgo(sigRes$res1M_WTG, ont = "CC"))

CCGO2F_WTG = summary(allgo(sigRes$res2F_WTG, ont = "CC"))
CCGO2M_WTG = summary(allgo(sigRes$res2M_WTG, ont = "CC"))

CCGO3F_WTG = summary(allgo(sigRes$res3F_WTG, ont = "CC"))
CCGO3M_WTG = summary(allgo(sigRes$res3M_WTG, ont = "CC"))


CCGO1F_WTG
CCGO1M_WTG
```

```{r}
CCGO2F_WTG
CCGO2M_WTG
```

```{r}
CCGO3F_WTG
CCGO3M_WTG
```

### MF PDE & NDE Analysis

```{r}
MFGO1F_WTG = summary(allgo(sigRes$res1F_WTG, ont = "MF"))
MFGO1M_WTG = summary(allgo(sigRes$res1M_WTG, ont = "MF"))

MFGO2F_WTG = summary(allgo(sigRes$res2F_WTG, ont = "MF"))
MFGO2M_WTG = summary(allgo(sigRes$res2M_WTG, ont = "MF"))

MFGO3F_WTG = summary(allgo(sigRes$res3F_WTG, ont = "MF"))
MFGO3M_WTG = summary(allgo(sigRes$res3M_WTG, ont = "MF"))

MFGO1F_WTG
MFGO1M_WTG
```

```{r}
MFGO2F_WTG
MFGO2M_WTG
```

```{r}
MFGO3F_WTG
MFGO3M_WTG
```

add comparisons 36vWT and 41vWT perhaps?

## Comparing different TG lines

### BP Analysis

```{r}
BPGOP1F_3641 = summary(go_PDE(sigRes$res1F_3641))
BPGOP1M_3641 = summary(go_PDE(sigRes$res1M_3641))
BPGON1F_3641 = summary(go_NDE(sigRes$res1F_3641))
BPGON1M_3641 = summary(go_NDE(sigRes$res1M_3641))

BPGOP2F_3641 = summary(go_PDE(sigRes$res2F_3641))
BPGOP2M_3641 = summary(go_PDE(sigRes$res2M_3641))
BPGON2F_3641 = summary(go_NDE(sigRes$res2F_3641))
BPGON2M_3641 = summary(go_NDE(sigRes$res2M_3641))

BPGOP3F_3641 = summary(go_PDE(sigRes$res3F_3641))
BPGOP3M_3641 = summary(go_PDE(sigRes$res3M_3641))
BPGON3F_3641 = summary(go_NDE(sigRes$res3F_3641))
BPGON3M_3641 = summary(go_NDE(sigRes$res3M_3641))

BPGOP1F_3641
BPGOP1M_3641
```

```{r}
BPGON1F_3641
BPGON1M_3641
```

```{r}
BPGOP2F_3641
BPGOP2M_3641
```

```{r}
BPGON2F_3641
BPGON2M_3641
```

```{r}
BPGOP3F_3641
BPGOP3M_3641
```

```{r}
BPGON3F_3641
BPGON3M_3641
```

### CC Analysis

```{r}
CCGOP1F_3641 = summary(go_PDE(sigRes$res1F_3641, ont = "CC"))
CCGOP1M_3641 = summary(go_PDE(sigRes$res1M_3641, ont = "CC"))
CCGON1F_3641 = summary(go_NDE(sigRes$res1F_3641, ont = "CC"))
CCGON1M_3641 = summary(go_NDE(sigRes$res1M_3641, ont = "CC"))

CCGOP2F_3641 = summary(go_PDE(sigRes$res2F_3641, ont = "CC"))
CCGOP2M_3641 = summary(go_PDE(sigRes$res2M_3641, ont = "CC"))
CCGON2F_3641 = summary(go_NDE(sigRes$res2F_3641, ont = "CC"))
CCGON2M_3641 = summary(go_NDE(sigRes$res2M_3641, ont = "CC"))

CCGOP3F_3641 = summary(go_PDE(sigRes$res3F_3641, ont = "CC"))
CCGOP3M_3641 = summary(go_PDE(sigRes$res3M_3641, ont = "CC"))
CCGON3F_3641 = summary(go_NDE(sigRes$res3F_3641, ont = "CC"))
CCGON3M_3641 = summary(go_NDE(sigRes$res3M_3641, ont = "CC"))

CCGOP1F_3641
CCGOP1M_3641
```

```{r}
CCGON1F_3641
CCGON1M_3641
```

```{r}
CCGOP2F_3641
CCGOP2M_3641
```

```{r}
CCGON2F_3641
CCGON2M_3641
```

```{r}
CCGOP3F_3641
CCGOP3M_3641
```

```{r}
CCGON3F_3641
CCGON3M_3641
```

### MF Analysis

```{r}
MFGOP1F_3641 = summary(go_PDE(sigRes$res1F_3641, ont = "MF"))
MFGOP1M_3641 = summary(go_PDE(sigRes$res1M_3641, ont = "MF"))
MFGON1F_3641 = summary(go_NDE(sigRes$res1F_3641, ont = "MF"))
MFGON1M_3641 = summary(go_NDE(sigRes$res1M_3641, ont = "MF"))

MFGOP2F_3641 = summary(go_PDE(sigRes$res2F_3641, ont = "MF"))
MFGOP2M_3641 = summary(go_PDE(sigRes$res2M_3641, ont = "MF"))
MFGON2F_3641 = summary(go_NDE(sigRes$res2F_3641, ont = "MF"))
MFGON2M_3641 = summary(go_NDE(sigRes$res2M_3641, ont = "MF"))

MFGOP3F_3641 = summary(go_PDE(sigRes$res3F_3641, ont = "MF"))
MFGOP3M_3641 = summary(go_PDE(sigRes$res3M_3641, ont = "MF"))
MFGON3F_3641 = summary(go_NDE(sigRes$res3F_3641, ont = "MF"))
MFGON3M_3641 = summary(go_NDE(sigRes$res3M_3641, ont = "MF"))

MFGOP1F_3641
MFGOP1M_3641
```

```{r}
MFGON1F_3641
MFGON1M_3641
```

```{r}
MFGOP2F_3641
MFGOP2M_3641
```

```{r}
MFGON2F_3641
MFGON2M_3641
```

```{r}
MFGOP3F_3641
MFGOP3M_3641
```

```{r}
MFGON3F_3641
MFGON3M_3641
```

### BP PDE & NDE Analysis

```{r}
BPGO1F_3641 = (summary(allgo(sigRes$res1F_3641)))
BPGO1M_3641 = (summary(allgo(sigRes$res1M_3641)))
BPGO2F_3641 = (summary(allgo(sigRes$res2F_3641)))
BPGO2M_3641 = (summary(allgo(sigRes$res2M_3641)))
BPGO3F_3641 = (summary(allgo(sigRes$res3F_3641)))
BPGO3M_3641 = (summary(allgo(sigRes$res3M_3641)))

BPGO1F_3641
BPGO1M_3641
```

```{r}
BPGO2F_3641
BPGO2M_3641
```

```{r}
BPGO3F_3641
BPGO3M_3641
```

### CC PDE & NDE Analysis

```{r}
CCGO1F_3641 = (summary(allgo(sigRes$res1F_3641, ont = "CC")))
CCGO1M_3641 = (summary(allgo(sigRes$res1M_3641, ont = "CC")))
CCGO2F_3641 = (summary(allgo(sigRes$res2F_3641, ont = "CC")))
CCGO2M_3641 = (summary(allgo(sigRes$res2M_3641, ont = "CC")))
CCGO3F_3641 = (summary(allgo(sigRes$res3F_3641, ont = "CC")))
CCGO3M_3641 = (summary(allgo(sigRes$res3M_3641, ont = "CC")))

CCGO1F_3641
CCGO1M_3641
```

```{r}
CCGO2F_3641
CCGO2M_3641
```

```{r}
CCGO3F_3641
CCGO3M_3641
```

### MF PDE & NDE Analysis

```{r}
MFGO1F_3641 = (summary(allgo(sigRes$res1F_3641, ont = "MF")))
MFGO1M_3641 = (summary(allgo(sigRes$res1M_3641, ont = "MF")))

MFGO2F_3641 = (summary(allgo(sigRes$res2F_3641, ont = "MF")))
MFGO2M_3641 = (summary(allgo(sigRes$res2M_3641, ont = "MF")))

MFGO3F_3641 = (summary(allgo(sigRes$res3F_3641, ont = "MF")))
MFGO3M_3641 = (summary(allgo(sigRes$res3M_3641, ont = "MF")))

MFGO1F_3641
MFGO1M_3641
```

```{r}
MFGO2F_3641
MFGO2M_3641
```

```{r}
MFGO3F_3641
MFGO3M_3641
```

## Timepoint Comparisons

### Line 36 BP Analysis

```{r}
BPGOPF_36_21 = (summary(go_PDE(sigRes$resF_36_21)))
BPGOPM_36_21 = (summary(go_PDE(sigRes$resM_36_21)))
BPGOMF_36_21 = (summary(go_NDE(sigRes$resF_36_21)))
BPGOMM_36_21 = (summary(go_NDE(sigRes$resM_36_21)))

BPGOPF_36_31 = (summary(go_PDE(sigRes$resF_36_31)))
BPGOPM_36_31 = (summary(go_PDE(sigRes$resM_36_31)))
BPGOMF_36_31 = (summary(go_NDE(sigRes$resF_36_31)))
BPGOMM_36_31 = (summary(go_NDE(sigRes$resM_36_31)))

BPGOPF_36_32 = (summary(go_PDE(sigRes$resF_36_32)))
BPGOPM_36_32 = (summary(go_PDE(sigRes$resM_36_32)))
BPGOMF_36_32 = (summary(go_NDE(sigRes$resF_36_32)))
BPGOMM_36_32 = (summary(go_NDE(sigRes$resM_36_32)))

BPGOPF_36_21
BPGOPM_36_21
```

```{r}
BPGOMF_36_21
BPGOMM_36_21
```

```{r}
BPGOPF_36_31
BPGOPM_36_31
```

```{r}
BPGOMF_36_31
BPGOMM_36_31
```

```{r}
BPGOPF_36_32
BPGOPM_36_32
```

```{r}
BPGOMF_36_32
BPGOMM_36_32
```

### Line 41 BP Analysis

```{r}
BPGOPF_41_21 = (summary(go_PDE(sigRes$resF_41_21)))
BPGOPM_41_21 = (summary(go_PDE(sigRes$resM_41_21)))
BPGOMF_41_21 = (summary(go_NDE(sigRes$resF_41_21)))
BPGOMM_41_21 = (summary(go_NDE(sigRes$resM_41_21)))

BPGOPF_41_31 = (summary(go_PDE(sigRes$resF_41_31)))
BPGOPM_41_31 = (summary(go_PDE(sigRes$resM_41_31)))
BPGOMF_41_31 = (summary(go_NDE(sigRes$resF_41_31)))
BPGOMM_41_31 = (summary(go_NDE(sigRes$resM_41_31)))

BPGOPF_41_32 = (summary(go_PDE(sigRes$resF_41_32)))
BPGOPM_41_32 = (summary(go_PDE(sigRes$resM_41_32)))
BPGOMF_41_32 = (summary(go_NDE(sigRes$resF_41_32)))
BPGOMM_41_32 = (summary(go_NDE(sigRes$resM_41_32)))

BPGOPF_41_21
BPGOPM_41_21
```

```{r}
BPGOMF_41_21
BPGOMM_41_21
```

```{r}
BPGOPF_41_31
BPGOPM_41_31
```

```{r}
BPGOMF_41_31
BPGOMM_41_31
```

```{r}
BPGOPF_41_32
BPGOPM_41_32
```

```{r}
BPGOMF_41_32
BPGOMM_41_32
```