---
title: "Visualizations"
author: "Lorenzo.MCM"
date: "2024-07-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Initial Setup

## Installing and loading packages.

```{r eval=FALSE}

install.packages(c('dplyr','readxl','DESeq2','ggplot2'), repos = "http://cran.us.r-project.org" )
```

```{r message=FALSE}
library(readxl)
library(DESeq2)
library(ggplot2)
library(gplots)
library(pheatmap)
library(vsn)
library(dplyr)
```

## Importing data and tidying up resulting data frame so rows are ID'd by Gene Symbol.NAs omitted for comparison with panther-db

```{r message=FALSE, warning=FALSE}

# Setting Working Directory (WD for easier output management)

#setwd("your/file/path")
setwd("C:/Users/Lorenzo/Documents/MESTRADO/R/RNAbps")

# Importing main reads file

data = as.data.frame(read_excel("Reads_genes.xlsx"))

# Keeping only the columns with MGI, RefSeq ID and Gene symbols + raw counts, omitting NAs 

reads = data %>% select(MGI, Gene_ID, Gene_Symbol, contains("Count"))

reads = as.data.frame(reads)

reads = na.omit(reads)

```

## Importing panther-db list of RBPs and tidying it up (browsed for RNA Metabolism Proteins and filtered for Mus musculus, exported to and edited in excel so the only row was MGI IDs, removed genes that did not have MGI IDs)

```{r message=FALSE, warning=FALSE}

# Importing list of RBPs, omitting NAs

mainPanther = as.data.frame(read_excel('Panther-RNA-ALL.xlsm', col_names = FALSE))

mainPanther = mainPanther[!is.na(mainPanther)]

# Filtering the reads object so it only contains genes whose MGI is also present in the list of RBPs, resulting in a final "rbps" object 
#(alternatively comment this first line and uncomment the last one to analyze all genes)

#comment this for analyzing all genes (uncomment last line)
rbps = reads[reads$MGI %in% mainPanther,]

#comment this for analyzing only rbps (uncomment previous line) 
#rbps = reads

```

## Naming conditions and final set up of rbps object

```{r message=FALSE, warning=FALSE}

# Naming rows in the rbps object after their MGI ID and removing all columns except those with the counts for each condition

rownames(rbps) = rbps$MGI

rbps = rbps[,-c(1,2,3)]

# Naming raw counts columns after their conditions, based on an external file

colnames(rbps) = (read_excel("conditions.xlsx", col_names = FALSE))[1,]
              
```

# Setting up for DESeq 
## Intermediate objects for selecting conditions for comparisons through DESeq

```{r message=FALSE, warning=FALSE}

#Splitting conditions for line and sex
rbpsF_WT = rbps %>% select(contains("FWT"))
rbpsM_WT = rbps %>% select(contains("MWT"))
rbpsF_TG = rbps %>% select(contains("F36"), contains("F41"))
rbpsM_TG = rbps %>% select(contains("M36"), contains("M41"))
rbpsF_36 = rbps %>% select(contains("F36"))
rbpsM_36 = rbps %>% select(contains("M36"))
rbpsF_41 = rbps %>% select(contains("F41"))
rbpsM_41 = rbps %>% select(contains("M41"))

#Final split into timepoints
T1F_41 = rbpsF_41[,grep("^1",names(rbpsF_41), value = TRUE)]
T1M_41 = rbpsM_41[,grep("^1",names(rbpsM_41), value = TRUE)]
T1F_WT = rbpsF_WT[,grep("^1",names(rbpsF_WT), value = TRUE)]
T1M_WT = rbpsM_WT[,grep("^1",names(rbpsM_WT), value = TRUE)]
T1F_36 = rbpsF_36[,grep("^1",names(rbpsF_36), value = TRUE)]
T1M_36 = rbpsM_36[,grep("^1",names(rbpsM_36), value = TRUE)]
T1F_TG = rbpsF_TG[,grep("^1",names(rbpsF_TG), value = TRUE)]
T1M_TG = rbpsM_TG[,grep("^1",names(rbpsM_TG), value = TRUE)]

T2F_41 = rbpsF_41[,grep("^2",names(rbpsF_41), value = TRUE)]
T2M_41 = rbpsM_41[,grep("^2",names(rbpsM_41), value = TRUE)]
T2F_WT = rbpsF_WT[,grep("^2",names(rbpsF_WT), value = TRUE)]
T2M_WT = rbpsM_WT[,grep("^2",names(rbpsM_WT), value = TRUE)]
T2F_36 = rbpsF_36[,grep("^2",names(rbpsF_36), value = TRUE)]
T2M_36 = rbpsM_36[,grep("^2",names(rbpsM_36), value = TRUE)]
T2F_TG = rbpsF_TG[,grep("^2",names(rbpsF_TG), value = TRUE)]
T2M_TG = rbpsM_TG[,grep("^2",names(rbpsM_TG), value = TRUE)]

T3F_41 = rbpsF_41[,grep("^3",names(rbpsF_41), value = TRUE)]
T3M_41 = rbpsM_41[,grep("^3",names(rbpsM_41), value = TRUE)]
T3F_WT = rbpsF_WT[,grep("^3",names(rbpsF_WT), value = TRUE)]
T3M_WT = rbpsM_WT[,grep("^3",names(rbpsM_WT), value = TRUE)]
T3F_36 = rbpsF_36[,grep("^3",names(rbpsF_36), value = TRUE)]
T3M_36 = rbpsM_36[,grep("^3",names(rbpsM_36), value = TRUE)]
T3F_TG = rbpsF_TG[,grep("^3",names(rbpsF_TG), value = TRUE)]
T3M_TG = rbpsM_TG[,grep("^3",names(rbpsM_TG), value = TRUE)]

#Binding final objects for comparison between different conditions (obj names = Timepoint-Sex-Line-conditions)
T1F_WT36 = bind_cols(T1F_WT, T1F_36)
T1M_WT36 = bind_cols(T1M_WT, T1M_36)
T1F_WT41 = bind_cols(T1F_WT, T1F_41)
T1M_WT41 = bind_cols(T1M_WT, T1M_41)
T1F_3641 = bind_cols(T1F_36, T1F_41)
T1M_3641 = bind_cols(T1M_36, T1M_41)
T1F_WTG = bind_cols(T1F_WT, T1F_TG)
T1M_WTG = bind_cols(T1M_WT, T1M_TG)
  
T2F_WT36 = bind_cols(T2F_WT, T2F_36)
T2M_WT36 = bind_cols(T2M_WT, T2M_36)
T2F_WT41 = bind_cols(T2F_WT, T2F_41)
T2M_WT41 = bind_cols(T2M_WT, T2M_41)
T2F_3641 = bind_cols(T2F_36, T2F_41)
T2M_3641 = bind_cols(T2M_36, T2M_41)
T2F_WTG = bind_cols(T2F_WT, T2F_TG)
T2M_WTG = bind_cols(T2M_WT, T2M_TG)

  
T3F_WT41 = bind_cols(T3F_WT, T3F_41)
T3M_WT41 = bind_cols(T3M_WT, T3M_41)
T3F_WT36 = bind_cols(T3F_WT, T3F_36)
T3M_WT36 = bind_cols(T3M_WT, T3M_36)
T3F_3641 = bind_cols(T3F_36, T3F_41)
T3M_3641 = bind_cols(T3M_36, T3M_41)
T3F_WTG = bind_cols(T3F_WT, T3F_TG)
T3M_WTG = bind_cols(T3M_WT, T3M_TG)

```

## setting up dfs for DESeq design 
```{r}
#timepoint lists
tpF_WT = c("1","1","1","2","2","2","3","3","3")
tpM_WT = c("1","1","1","2","2","2","3","3","3")
tpF_TG = c("1","1","1","1","2","2","2","3","1","1","1","2","2","2","3","3")
tpM_TG = c("1","1","2","2","2","3","3","3","3","3","1","1","1","2","2","2","3","3","3","3")

tpF_36 = c("1","1","1","1","2","2","2","3")
tpM_36 = c("1","1","2","2","2","3","3","3","3","3")

tpF_41 = c("1","1","1","2","2","2","3","3")
tpM_41 = c("1","1","1","2","2","2","3","3","3","3")

#line list

l1F_3641 = c("36","36","36","36","41","41","41")
l1M_3641 = c("36","36","41","41","41")

l1F_WTG = c("WT","WT","WT","TG","TG","TG","TG","TG","TG","TG")
l1M_WTG = c("WT","WT","WT","TG","TG","TG","TG","TG")

l1F_WT36 = c("WT","WT","WT","36","36","36","36")
l1M_WT36 = c("WT","WT","WT","36","36")

l1F_WT41 = c("WT","WT","WT","41","41","41")
l1M_WT41 = c("WT","WT","WT","41","41","41")

l2F_WTG = c("WT","WT","WT","TG","TG","TG","TG","TG","TG")
l2M_WTG = c("WT","WT","WT","TG","TG","TG","TG","TG","TG")

l2F_WT36 = c("WT","WT","WT","36","36","36")
l2M_WT36 = c("WT","WT","WT","36","36","36")

l2F_WT41 = c("WT","WT","WT","41","41","41")
l2M_WT41 = c("WT","WT","WT","41","41","41")

l2F_3641 = c("36","36","36","41","41","41")
l2M_3641 = c("36","36","36","41","41","41")

l3F_WTG = c("WT","WT","WT","TG","TG","TG")
l3M_WTG = c("WT","WT","WT","TG","TG","TG","TG","TG","TG","TG","TG","TG")

l3F_WT36 = c("WT","WT","WT","36")
l3M_WT36 = c("WT","WT","WT","36","36","36","36","36")

l3F_WT41 = c("WT","WT","WT","41","41")
l3M_WT41 = c("WT","WT","WT","41","41","41","41")

l3F_3641 = c("36","41","41")
l3M_3641 = c("36","36","36","36","36","41","41","41", "41")

# Binding timepoints and line lists for each condition

c1F_WTG = as.data.frame(cbind(colnames(T1F_WTG),l1F_WTG))
c1M_WTG = as.data.frame(cbind(colnames(T1M_WTG),l1M_WTG))

c1F_WT36 = as.data.frame(cbind(colnames(T1F_WT36),l1F_WT36))
c1M_WT36 = as.data.frame(cbind(colnames(T1M_WT36),l1M_WT36))

c1F_WT41 = as.data.frame(cbind(colnames(T1F_WT41),l1F_WT41))
c1M_WT41 = as.data.frame(cbind(colnames(T1M_WT41),l1M_WT41))

c1F_3641 = as.data.frame(cbind(colnames(T1F_3641),l1F_3641))
c1M_3641 = as.data.frame(cbind(colnames(T1M_3641),l1M_3641))

c2F_WTG = as.data.frame(cbind(colnames(T2F_WTG),l2F_WTG))
c2M_WTG = as.data.frame(cbind(colnames(T2M_WTG),l2M_WTG))

c2F_WT36 = as.data.frame(cbind(colnames(T2F_WT36),l2F_WT36))
c2M_WT36 = as.data.frame(cbind(colnames(T2M_WT36),l2M_WT36))

c2F_WT41 = as.data.frame(cbind(colnames(T2F_WT41),l2F_WT41))
c2M_WT41 = as.data.frame(cbind(colnames(T2M_WT41),l2M_WT41))

c2F_3641 = as.data.frame(cbind(colnames(T2F_3641),l2F_3641))
c2M_3641 = as.data.frame(cbind(colnames(T2M_3641),l2M_3641))

c3F_WTG = as.data.frame(cbind(colnames(T3F_WTG),l3F_WTG))
c3M_WTG = as.data.frame(cbind(colnames(T3M_WTG),l3M_WTG))

c3F_WT36 = as.data.frame(cbind(colnames(T3F_WT36),l3F_WT36))
c3M_WT36 = as.data.frame(cbind(colnames(T3M_WT36),l3M_WT36))

c3F_WT41 = as.data.frame(cbind(colnames(T3F_WT41),l3F_WT41))
c3M_WT41 = as.data.frame(cbind(colnames(T3M_WT41),l3M_WT41))

c3F_3641 = as.data.frame(cbind(colnames(T3F_3641),l3F_3641))
c3M_3641 = as.data.frame(cbind(colnames(T3M_3641),l3M_3641))

cF_WT = as.data.frame(cbind(colnames(rbpsF_WT),tpF_WT))
cM_WT = as.data.frame(cbind(colnames(rbpsM_WT),tpM_WT))

cF_TG = as.data.frame(cbind(colnames(rbpsF_TG),tpF_TG))
cM_TG = as.data.frame(cbind(colnames(rbpsM_TG),tpM_TG))

cF_36 = as.data.frame(cbind(colnames(rbpsF_36),tpF_36))
cM_36 = as.data.frame(cbind(colnames(rbpsM_36),tpM_36))

cF_41 = as.data.frame(cbind(colnames(rbpsF_41), tpF_41))
cM_41 = as.data.frame(cbind(colnames(rbpsM_41), tpM_41))


```

## Conversion into DESeqDataSets

```{r message=FALSE, warning=FALSE, include=FALSE}

dds1F_WTG = DESeqDataSetFromMatrix(countData = T1F_WTG,colData = c1F_WTG,
                                  design = ~l1F_WTG)
dds1M_WTG = DESeqDataSetFromMatrix(countData = T1M_WTG,colData = c1M_WTG,
                                  design = ~l1M_WTG)

dds1F_WT36 = DESeqDataSetFromMatrix(countData = T1F_WT36, colData = c1F_WT36,
                                 design = ~l1F_WT36)
dds1M_WT36 = DESeqDataSetFromMatrix(countData = T1M_WT36, colData = c1M_WT36,
                                 design = ~l1M_WT36)

dds1F_WT41 = DESeqDataSetFromMatrix(countData = T1F_WT41,colData = c1F_WT41,
                                 design = ~l1F_WT41)
dds1M_WT41 = DESeqDataSetFromMatrix(countData = T1M_WT41,colData = c1M_WT41,
                                 design = ~l1M_WT41)

dds1F_3641 = DESeqDataSetFromMatrix(countData = T1F_3641,colData = c1F_3641,
                                    design = ~l1F_3641)
dds1M_3641 = DESeqDataSetFromMatrix(countData = T1M_3641,colData = c1M_3641,
                                    design = ~l1M_3641)

dds2F_WTG = DESeqDataSetFromMatrix(countData = T2F_WTG,colData = c2F_WTG,
                                  design = ~l2F_WTG)
dds2M_WTG = DESeqDataSetFromMatrix(countData = T2M_WTG,colData = c2M_WTG,
                                  design = ~l2M_WTG)

dds2F_WT36 = DESeqDataSetFromMatrix(countData = T2F_WT36,colData = c2F_WT36,
                                 design = ~l2F_WT36)
dds2M_WT36 = DESeqDataSetFromMatrix(countData = T2M_WT36,colData = c2M_WT36,
                                 design = ~l2M_WT36)

dds2F_WT41 = DESeqDataSetFromMatrix(countData = T2F_WT41,colData = c2F_WT41,
                                 design = ~l2F_WT41)
dds2M_WT41 = DESeqDataSetFromMatrix(countData = T2M_WT41,colData = c2M_WT41,
                                 design = ~l2M_WT41)

dds2F_3641 = DESeqDataSetFromMatrix(countData = T2F_3641,colData = c2F_3641,
                                    design = ~l2F_3641)
dds2M_3641 = DESeqDataSetFromMatrix(countData = T2M_3641,colData = c2M_3641,
                                    design = ~l2M_3641)

dds3F_WTG = DESeqDataSetFromMatrix(countData = T3F_WTG,colData = c3F_WTG,
                                  design = ~l3F_WTG)
dds3M_WTG = DESeqDataSetFromMatrix(countData = T3M_WTG,colData = c3M_WTG,
                                  design = ~l3M_WTG)

dds3F_WT36 = DESeqDataSetFromMatrix(countData = T3F_WT36,colData = c3F_WT36,
                                 design = ~l3F_WT36)
dds3M_WT36 = DESeqDataSetFromMatrix(countData = T3M_WT36,colData = c3M_WT36,
                                 design = ~l3M_WT36)

dds3F_WT41 = DESeqDataSetFromMatrix(countData = T3F_WT41,colData = c3F_WT41,
                                 design = ~l3F_WT41)
dds3M_WT41 = DESeqDataSetFromMatrix(countData = T3M_WT41, colData = c3M_WT41,
                                 design = ~l3M_WT41)

dds3F_3641 = DESeqDataSetFromMatrix(countData = T3F_3641,colData = c3F_3641,
                                    design = ~l3F_3641)
dds3M_3641 = DESeqDataSetFromMatrix(countData = T3M_3641,colData = c3M_3641,
                                    design = ~l3M_3641)

ddsF_WT = DESeqDataSetFromMatrix(countData = rbpsF_WT,colData = cF_WT,
                                design = ~tpF_WT)
ddsM_WT = DESeqDataSetFromMatrix(countData = rbpsM_WT,colData = cM_WT,
                                design = ~tpM_WT)

ddsF_TG = DESeqDataSetFromMatrix(countData = rbpsF_TG,colData = cF_TG,
                                design = ~tpF_TG)
ddsM_TG = DESeqDataSetFromMatrix(countData = rbpsM_TG,colData = cM_TG,
                                design = ~tpM_TG)

ddsF_36 = DESeqDataSetFromMatrix(countData = rbpsF_36,colData = cF_36,
                                design = ~tpF_36)
ddsM_36 = DESeqDataSetFromMatrix(countData = rbpsM_36,colData = cM_36,
                                design = ~tpM_36)

ddsF_41 = DESeqDataSetFromMatrix(countData = rbpsF_41,colData = cF_41,
                                design = ~tpF_41)
ddsM_41 = DESeqDataSetFromMatrix(countData = rbpsM_41,colData = cM_41,
                                design = ~tpM_41)
```

## Running DESeq

```{r message=FALSE, warning=FALSE}

dds1F_WTG = DESeq(dds1F_WTG)
dds1M_WTG = DESeq(dds1M_WTG)

dds1F_WT36 = DESeq(dds1F_WT36)
dds1M_WT36 = DESeq(dds1M_WT36)

dds1F_WT41 = DESeq(dds1F_WT41)
dds1M_WT41 = DESeq(dds1M_WT41)

dds1F_3641 = DESeq(dds1F_3641)
dds1M_3641 = DESeq(dds1M_3641)

dds2F_WTG = DESeq(dds2F_WTG)
dds2M_WTG = DESeq(dds2M_WTG)

dds2F_WT36 = DESeq(dds2F_WT36)
dds2M_WT36 = DESeq(dds2M_WT36)

dds2F_WT41 = DESeq(dds2F_WT41)
dds2M_WT41 = DESeq(dds2M_WT41)

dds2F_3641 = DESeq(dds2F_3641)
dds2M_3641 = DESeq(dds2M_3641)

dds3F_WTG = DESeq(dds3F_WTG)
dds3M_WTG = DESeq(dds3M_WTG)

dds3F_WT36 = DESeq(dds3F_WT36)
dds3M_WT36 = DESeq(dds3M_WT36)

dds3F_WT41 = DESeq(dds3F_WT41)
dds3M_WT41 = DESeq(dds3M_WT41)

dds3F_3641 = DESeq(dds3F_3641)
dds3M_3641 = DESeq(dds3M_3641)

ddsF_WT = DESeq(ddsF_WT)
ddsM_WT = DESeq(ddsM_WT)

ddsF_TG = DESeq(ddsF_TG)
ddsM_TG = DESeq(ddsM_TG)

ddsF_36 = DESeq(ddsF_36)
ddsM_36 = DESeq(ddsM_36)

ddsF_41 = DESeq(ddsF_41)
ddsM_41 = DESeq(ddsM_41)
```
```{r}

## Filtering for genes read less than 5 times across all samples (Only for histogram visualizations)


dds1F_WTG = dds1F_WTG[rowSums(counts(dds1F_WTG)) > 5,]
dds1M_WTG = dds1M_WTG[rowSums(counts(dds1M_WTG)) > 5,]

dds1F_WT36 = dds1F_WT36[rowSums(counts(dds1F_WT36)) > 5,]
dds1M_WT36 = dds1M_WT36[rowSums(counts(dds1M_WT36)) > 5,]

dds1F_WT41 = dds1F_WT41[rowSums(counts(dds1F_WT41)) > 5,]
dds1M_WT41 = dds1M_WT41[rowSums(counts(dds1M_WT41)) > 5,]

dds1F_3641 = dds1F_3641[rowSums(counts(dds1F_3641)) > 5,]
dds1M_3641 = dds1M_3641[rowSums(counts(dds1M_3641)) > 5,]

dds2F_WTG = dds2F_WTG[rowSums(counts(dds2F_WTG)) > 5,]
dds2M_WTG = dds2M_WTG[rowSums(counts(dds2M_WTG)) > 5,]

dds2F_WT36 = dds2F_WT36[rowSums(counts(dds2F_WT36)) > 5,]
dds2M_WT36 = dds2M_WT36[rowSums(counts(dds2M_WT36)) > 5,]

dds2F_WT41 = dds2F_WT41[rowSums(counts(dds2F_WT41)) > 5,]
dds2M_WT41 = dds2M_WT41[rowSums(counts(dds2M_WT41)) > 5,]

dds2F_3641 = dds2F_3641[rowSums(counts(dds2F_3641)) > 5,]
dds2M_3641 = dds2M_3641[rowSums(counts(dds2M_3641)) > 5,]

dds3F_WTG = dds3F_WTG[rowSums(counts(dds3F_WTG)) > 5,]
dds3M_WTG = dds3M_WTG[rowSums(counts(dds3M_WTG)) > 5,]

dds3F_WT36 = dds3F_WT36[rowSums(counts(dds3F_WT36)) > 5,]
dds3M_WT36 = dds3M_WT36[rowSums(counts(dds3M_WT36)) > 5,]

dds3F_WT41 = dds3F_WT41[rowSums(counts(dds3F_WT41)) > 5,]
dds3M_WT41 = dds3M_WT41[rowSums(counts(dds3M_WT41)) > 5,]

dds3F_3641 = dds3F_3641[rowSums(counts(dds3F_3641)) > 5,]
dds3M_3641 = dds3M_3641[rowSums(counts(dds3M_3641)) > 5,]

ddsF_WT = ddsF_WT[rowSums(counts(ddsF_WT)) > 5,]
ddsM_WT = ddsM_WT[rowSums(counts(ddsM_WT)) > 5,]

ddsF_TG = ddsF_TG[rowSums(counts(ddsF_TG)) > 5,]
ddsM_TG = ddsM_TG[rowSums(counts(ddsM_TG)) > 5,]

ddsF_36 = ddsF_36[rowSums(counts(ddsF_36)) > 5,]
ddsM_36 = ddsM_36[rowSums(counts(ddsM_36)) > 5,]

ddsF_41 = ddsF_41[rowSums(counts(ddsF_41)) > 5,]
ddsM_41 = ddsM_41[rowSums(counts(ddsM_41)) > 5,]

```

```{r}

## log 2 transform 
rl1F_WTG = rlog(dds1F_WTG)
rl1M_WTG = rlog(dds1M_WTG)

rl1F_WT36 = rlog(dds1F_WT36)
rl1M_WT36 = rlog(dds1M_WT36)

rl1F_WT41 = rlog(dds1F_WT41)
rl1M_WT41 = rlog(dds1M_WT41)

rl1F_3641 = rlog(dds1F_3641)
rl1M_3641 = rlog(dds1M_3641)

rl2F_WTG = rlog(dds2F_WTG)
rl2M_WTG = rlog(dds2M_WTG)

rl2F_WT36 = rlog(dds2F_WT36)
rl2M_WT36 = rlog(dds2M_WT36)

rl2F_WT41 = rlog(dds2F_WT41)
rl2M_WT41 = rlog(dds2M_WT41)

rl2F_3641 = rlog(dds2F_3641)
rl2M_3641 = rlog(dds2M_3641)

rl3F_WTG = rlog(dds3F_WTG)
rl3M_WTG = rlog(dds3M_WTG)

rl3F_WT36 = rlog(dds3F_WT36)
rl3M_WT36 = rlog(dds3M_WT36)

rl3F_WT41 = rlog(dds3F_WT41)
rl3M_WT41 = rlog(dds3M_WT41)

rl3F_3641 = rlog(dds3F_3641)
rl3M_3641 = rlog(dds3M_3641)

rlF_WT = rlog(ddsF_WT)
rlM_WT = rlog(ddsM_WT)

rlF_TG = rlog(ddsF_TG)
rlM_TG = rlog(ddsM_TG)

rlF_36 = rlog(ddsF_36)
rlM_36 = rlog(ddsM_36)

rlF_41 = rlog(ddsF_41)
rlM_41 = rlog(ddsM_41)

```

```{r}
plotPCA(rl1F_WTG, intgroup= c("l1F_WTG"))
plotPCA(rl1M_WTG, intgroup= c("l1M_WTG"))

plotPCA(rl1F_WT36, intgroup = c("l1F_WT36"))
plotPCA(rl1M_WT36, intgroup = c("l1M_WT36"))

plotPCA(rl1F_WT41, intgroup = c("l1F_WT41"))
plotPCA(rl1M_WT41, intgroup = c("l1M_WT41"))

plotPCA(rl1F_3641, intgroup = c("l1F_3641"))
plotPCA(rl1M_3641, intgroup = c("l1M_3641"))
```

```{r}
plotPCA(rl2F_WTG, intgroup= c("l2F_WTG"))
plotPCA(rl2M_WTG, intgroup= c("l2M_WTG"))

plotPCA(rl2F_WT36, intgroup = c("l2F_WT36"))
plotPCA(rl2M_WT36, intgroup = c("l2M_WT36"))

plotPCA(rl2F_WT41, intgroup = c("l2F_WT41"))
plotPCA(rl2M_WT41, intgroup = c("l2M_WT41"))

plotPCA(rl2F_3641, intgroup = c("l2F_3641"))
plotPCA(rl2M_3641, intgroup = c("l2M_3641"))
```

```{r}
plotPCA(rl3F_WTG, intgroup= c("l3F_WTG"))
plotPCA(rl3M_WTG, intgroup= c("l3M_WTG"))

plotPCA(rl3F_WT36, intgroup = c("l3F_WT36"))
plotPCA(rl3M_WT36, intgroup = c("l3M_WT36"))

plotPCA(rl3F_WT41, intgroup = c("l3F_WT41"))
plotPCA(rl3M_WT41, intgroup = c("l3M_WT41"))

plotPCA(rl3F_3641, intgroup = c("l3F_3641"))
plotPCA(rl3M_3641, intgroup = c("l3M_3641"))
```

```{r}
plotPCA(rlF_WT, intgroup = c("tpF_WT"))
plotPCA(rlM_WT, intgroup = c("tpM_WT"))

plotPCA(rlF_36, intgroup = c("tpF_36"))
plotPCA(rlM_36, intgroup = c("tpM_36"))

plotPCA(rlF_41, intgroup = c("tpF_41"))
plotPCA(rlM_41, intgroup = c("tpM_41"))

plotPCA(rlF_TG, intgroup = c("tpF_TG"))
plotPCA(rlM_TG, intgroup = c("tpM_TG"))

```