---
title: "WGCNA"
author: "Lorenzo.MCM"
date: "2024-06-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Initial Setup

## Installing and loading packages.

```{r eval=FALSE}

install.packages(c('dplyr','biomaRt','readxl','DESeq2'), repos = "http://cran.us.r-project.org" )
```

```{r eval=FALSE}
BiocManager::install(c('DESeq2', 'AnnotationDbi', 'org.Mm.eg.db', 'apeglm', 'ReportingTools', 'clusterProfiler'))

install.packages("remotes")
remotes::install_github("kevinblighe/CorLevelPlot")
```

```{r message=FALSE}
library(readxl)
library(DESeq2)
library(stringr)
library(ggplot2)
library(dplyr)
library(CorLevelPlot)
```

## Importing data and tidying up resulting data frame so rows are ID'd by Gene Symbol.NAs omitted for comparison with Gene list

```{r message=FALSE, warning=FALSE}

# Setting Working Directory 

#setwd("your/file/path")
setwd("C:/Users/Lorenzo/Documents/MESTRADO/R/RNAbps")

# Importing main reads file

data = as.data.frame(read_excel("Reads_genes.xlsx"))

# Keeping only the columns with MGI, RefSeq ID and Gene symbols + raw counts, omitting NAs 

reads = data %>% select(MGI, Gene_ID, Gene_Symbol, contains("Count"))

reads = as.data.frame(reads)

reads = na.omit(reads)

rbps = reads

```

## Naming conditions and final set up of rbps object

```{r message=FALSE, warning=FALSE}

# Naming rows in the rbps object after their MGI ID and removing all columns except those with the counts for each condition

rownames(rbps) = rbps$Gene_Symbol

rbps = rbps[,-c(1,2,3)]

# Naming raw counts columns after their conditions, based on an external file

colnames(rbps) = (read_excel("conditions.xlsx", col_names = FALSE))[1,]


# For analyzing c41
rbps = rbps[,c(1:18, 37:54)]
```

# Allowing multithreading and checking data for outlier samples and genes

```{r}
library(WGCNA)
allowWGCNAThreads()
gsg = t(rbps)

gsg = goodSamplesGenes(t(rbps))

```

# Removing outlier genes

```{r}
rbps = rbps[gsg$goodGenes == TRUE,]
```


## setting up dfs for DESeq design 
```{r}

#for 41

tp = c("1", "1", "1", "1", "1", "1", "2", "2", "2", "2", "2", "2", "3", "3", "3", "3", "3", "3", "1", "1", "1", "1", "1", "1", "2", "2", "2", "2", "2", "2", "3", "3", "3", "3", "3", "3")

colony = c("WT", "WT", "WT", "WT", "WT", "WT", "WT", "WT", "WT", "WT", "WT", "WT", "WT", "WT", "WT", "WT", "WT", "WT", "41", "41", "41", "41", "41", "41", "41", "41", "41", "41", "41", "41", "41", "41", "41", "41", "41", "41")

sex = c("F", "F", "F", "M", "M", "M", "F", "F", "F", "M", "M", "M", "F", "F", "F", "M", "M", "M", "F", "F", "F", "M", "M", "M", "F", "F", "F", "M", "M", "M", "F", "F", "M", "M", "M", "M")


nameCond = (read_excel("conditions.xlsx", col_names = FALSE))

nameCond = nameCond[,c(1:18, 37:54)] 
  
columnData = as.data.frame(cbind(rownames(nameCond[,1]),colony, tp, sex))
columnData = columnData[,-1]
rownames(columnData) = nameCond[1,]
```

## Conversion into DESeqDataSets

```{r message=FALSE, warning=FALSE, include=FALSE}

dds = DESeqDataSetFromMatrix(countData = rbps,
                            colData = columnData,
                            design = ~1)

```

## Filtering dataset for those genes which have at least 15 counts in 75% of conditions

```{r}
dds = dds[rowSums(counts(dds) >= 15) >= 27]
```

```{r}
## variance stabilization
dds_norm = vst(dds)
```

```{r}
norm.counts = assay(dds_norm) %>% t()
```

## Determining a soft power threshold(how strong connections must be for determining modules)

```{r}
power = c(c(1:10), seq(from = 12, to = 30, by = 2))
```

```{r}
sft = pickSoftThreshold(norm.counts,
                        powerVector = power,
                        networkType = 'signed',
                        verbose = 5)

sft.data = sft$fitIndices
```

## Ideally we want to pick something with a model fit above 0.8 and not too low of a mean connectivity, in this case I chose 26

```{r}
a1 = ggplot(sft.data, aes(Power, SFT.R.sq, label = Power))+
  geom_point() +
  geom_text(nudge_y = 0.1) +
  geom_hline(yintercept = 0.8, color = 'red') +
  labs(x = 'Power', y = 'Scale free topology model fit, signed R^2') +
  theme_classic()

a2 = ggplot(sft.data, aes(Power,mean.k., label = Power))+
  geom_point() +
  geom_text(nudge_y = 0.1)+
  labs(x= 'Power', y = 'Mean Connectivity') +
  theme_classic()

library(gridExtra)

grid.arrange(a1, a2, nrow = 2)
```

```{r}
norm.counts[] = sapply(norm.counts, as.numeric)

soft_power = 26

# This has a tendency to not work if I don't keep this in
temp_cor = cor
cor = WGCNA::cor
```

## Separating dataset into modules (takes about 15-20 minutes on my CPU)

```{r}

bwnet = blockwiseModules(norm.counts,
                 maxBlockSize = 16000,
                 TOMType = 'signed',
                 power = soft_power,
                 mergeCutHeight = 0.25,
                 numericLabels = FALSE,
                 randomSeed = 1234,
                 verbose = 3)

```

```{r}
cor = temp_cor
module_eigengenes = bwnet$MEs
```

```{r, fig.width=10}
plotDendroAndColors(bwnet$dendrograms[[1]], cbind(bwnet$unmergedColors,bwnet$colors),
                    c('unmerged', 'merged'),
                    dendroLabels = FALSE,
                    addGuide = TRUE,
                    hang = 0.03,
                    guideHang = 0.05)
```

```{r}
traits = columnData %>% 
  mutate(sex_bin = ifelse(grepl('F', sex), 1, 0))
```

```{r}

traits1 = columnData %>% 
  mutate(c41_bin = ifelse(grepl('41', colony), 1, 0))


traits$c41_bin = traits1$c41_bin

traits1 = columnData %>% 
  mutate(tp1 = ifelse(grepl('1', tp), 1, 0))

traits$tp1_bin = traits1$tp1

traits1 = columnData %>% 
  mutate(tp2 = ifelse(grepl('2', tp), 1, 0))

traits$tp2_bin = traits1$tp2

traits1 = columnData %>% 
  mutate(tp3 = ifelse(grepl('3', tp), 1, 0))

traits$tp3_bin = traits1$tp3

traits$F1_41 = ifelse(traits$sex_bin == 1 & traits$c41_bin == 1 & traits$tp1_bin == 1, 1, 0)
traits$F2_41 = ifelse(traits$sex_bin == 1 & traits$c41_bin == 1 & traits$tp2_bin == 1, 1, 0)
traits$F3_41 = ifelse(traits$sex_bin == 1 & traits$c41_bin == 1 & traits$tp3_bin == 1, 1, 0)

traits$M1_41 = ifelse(traits$sex_bin == 0 & traits$c41_bin == 1 & traits$tp1_bin == 1, 1, 0)
traits$M2_41 = ifelse(traits$sex_bin == 0 & traits$c41_bin == 1 & traits$tp2_bin == 1, 1, 0)
traits$M3_41 = ifelse(traits$sex_bin == 0 & traits$c41_bin == 1 & traits$tp3_bin == 1, 1, 0)


rm(traits1)
```

```{r}
nSamples = nrow(norm.counts)
nGenes = ncol(norm.counts)

module.trait.corr = cor(module_eigengenes, traits, use = 'p')
module.trait.pvals = corPvalueStudent(module.trait.corr, nSamples)
```

```{r, fig.width= 15}
heatmap.data = merge(module_eigengenes, traits, by = 'row.names')

CorLevelPlot(heatmap.data,
             x = names(heatmap.data)[c(19:20,24:29)],
             y = names(heatmap.data) [2:15],
             col = c('blue1', 'skyblue', 'white', 'pink', 'red'))
```

## Hub Genes sorted modules

```{r}
moduleColors = bwnet$colors

moduleColorsDF = as.data.frame(moduleColors)

geneTree = bwnet$dendrograms[[1]]
```

```{r}
adjacency_matrix = adjacency(norm.counts, power = soft_power)

TOM = TOMsimilarity(adjacency_matrix)
```

```{r}
kWithin = intramodularConnectivity(adjacency_matrix, moduleColors)
```

```{r}
hubGenes = function(color, top = NULL){
  
  modules_genes = which(moduleColors == color)
  
  kWithin_module = kWithin[modules_genes,]
  
  kWithin_sorted = kWithin_module[order(-kWithin_module$kWithin),]
  
  if(is.null(top) == FALSE)
  {return(as.data.frame(rownames(kWithin_sorted)[1:top]))}
  else(return(as.data.frame(rownames(kWithin_sorted))))
}
```

### Exporting list of genes for each module sorted by intramodular connectivity, so the first ones are Hub genes

```{r}
write.csv(hubGenes(color = 'brown'), file = "brownMod41.csv")

write.csv(hubGenes(color = 'grey'), file = "greyMod41.csv")

write.csv(hubGenes(color = 'green'), file = "greenMod41.csv")

write.csv(hubGenes(color = 'black'), file = "blackMod41.csv")

write.csv(hubGenes(color = 'blue'), file = "blueMod41.csv")

write.csv(hubGenes(color = 'salmon'), file = "salmonMod41.csv")

write.csv(hubGenes(color = 'red'), file = "redMod41.csv")

write.csv(hubGenes(color = 'greenyellow'), file = "greenyellowMod41.csv")

write.csv(hubGenes(color = 'turquoise'), file = "turquoiseMod41.csv")

write.csv(hubGenes(color = 'pink'), file = "pinkMod41.csv")

write.csv(hubGenes(color = 'magenta'), file = "magentaMod41.csv")

write.csv(hubGenes(color = 'yellow'), file = "yellowMod41.csv")

write.csv(hubGenes(color = 'tan'), file = "tanMod41.csv")

write.csv(hubGenes(color = 'purple'), file = "purpleMod41.csv")
```

### Function for outputing cytoscape files for specific modules

```{r}
colornw = function(x){
  
  moduleOfInterest = x
  module_genes = which(moduleColors == moduleOfInterest)
  module_gene_names = colnames(norm.counts)[module_genes]
  
  TOM_module = TOM[module_genes,module_genes]
  
  exportNetworkToCytoscape(
    TOM_module,
    edgeFile = paste("CSEdges_", moduleOfInterest, ".txt", sep = ""),
    nodeFile = paste("CSNodes_", moduleOfInterest, ".txt", sep = ""),
    threshold = 0.02,
    nodeNames = module_gene_names,
    nodeAttr = moduleColors[module_genes]
  )
  
}

```

